---
title: "SNAP_RCP_DataExtractionOrganization"
author: "Amaryllis Adey"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setting up the workspace

```{r}
library(dplyr)
library(ggplot2)
```

# Attempting to extract the RCP data from the .csv file

## Starting with the RCP 4.5 data

### Extract the data

#### Copper River Delta

```{r}
CR_snapTemp <- read.csv("DataFiles/CopperRiver_CreatedFiles/CordovaMonthlyTemperatures.csv", header = TRUE)

# Pull out only the RCP 4.5 data
CR_4.5 <- CR_snapTemp %>%
  filter(scenario == "rcp45")

# Check the model options here
unique(CR_4.5$model) # "5ModelAvg"  "GFDL-CM3"   "NCAR-CCSM4"

# using the 5ModelAvg for now 
CR_4.5 <- CR_snapTemp %>%
  filter(scenario == "rcp45" & model == "5ModelAvg")

# reorder by month-date
test <- CR_4.5 %>%
  dplyr::mutate(date = as.Date(date, "%Y%m%d")) %>%
  dplyr::select(date, tasmin, tasmean, tasmax) %>%
  dplyr::arrange(date)

test$std <- (test$tasmax - test$tasmin)/2
test$precision <- (1/(test$std^2))

CR_4.5 <- test
CR_4.5
```

#### Yakutat Forelands

```{r}
YF_snapTemp <- read.csv("DataFiles/YakutatForelands_CreatedFiles/YakutatMonthlyTemperatures.csv", header = TRUE)

# Pull out only the RCP 4.5 data
YF_4.5 <- YF_snapTemp %>%
  filter(scenario == "rcp45")

# Check the model options here
unique(YF_4.5$model) # "5ModelAvg"  "GFDL-CM3"   "NCAR-CCSM4"

# using the 5ModelAvg for now 
YF_4.5 <- YF_snapTemp %>%
  filter(scenario == "rcp45" & model == "5ModelAvg")

# reorder by month-date
test <- YF_4.5 %>%
  dplyr::mutate(date = as.Date(date, "%Y%m%d")) %>%
  dplyr::select(date, tasmin, tasmean, tasmax) %>%
  dplyr::arrange(date)

test$std <- (test$tasmax - test$tasmin)/2
test$precision <- (1/(test$std^2))

YF_4.5 <- test
YF_4.5
```

### Plot the forecasted data

#### Copper River Delta

```{r}
CR_4.5

# Full time frame (2006-2100)
CR_SNAP <- ggplot(data = CR_4.5, mapping = aes(x = date)) +
  geom_line(aes(y = tasmean), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Copper River SNAP Forecast - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
CR_SNAP

ggsave("CR_SNAP_Forecast4.5.jpeg", plot = CR_SNAP, width = 8, height = 6)

# Limited time frame (2020-2100)
test <- CR_4.5 %>%
  filter(date > "2019-12-01")

CR_4.5 <- test

CR_SNAP <- ggplot(data = CR_4.5, mapping = aes(x = date)) +
  geom_line(aes(y = tasmean), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Copper River SNAP Forecast - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
CR_SNAP

ggsave("CR_SNAP_Forecast4.5_2020.jpeg", plot = CR_SNAP, width = 8, height = 6)
```

#### Yakutat Forelands

```{r}
YF_4.5

# Full time frame (2006-2100)
YF_SNAP <- ggplot(data = YF_4.5, mapping = aes(x = date)) +
  geom_line(aes(y = tasmean), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Yakutat Forelands SNAP Forecast - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
YF_SNAP

ggsave("YF_SNAP_Forecast4.5.jpeg", plot = YF_SNAP, width = 8, height = 6)

# Limited time frame (2020-2100)
test <- YF_4.5 %>%
  filter(date > "2019-12-01")

YF_4.5 <- test

YF_SNAP <- ggplot(data = YF_4.5, mapping = aes(x = date)) +
  geom_line(aes(y = tasmean), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Yakutat Forelands SNAP Forecast - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
YF_SNAP

ggsave("YF_SNAP_Forecast4.5_2020.jpeg", plot = YF_SNAP, width = 8, height = 6)
```

## Continuing with the RCP 8.5 data

### Extract the data

#### Copper River Delta

```{r}
CR_snapTemp <- read.csv("DataFiles/CopperRiver_CreatedFiles/CordovaMonthlyTemperatures.csv", header = TRUE)

# Pull out only the RCP 4.5 data
CR_8.5 <- CR_snapTemp %>%
  filter(scenario == "rcp85")

# Check the model options here
unique(CR_8.5$model) # "5ModelAvg"  "GFDL-CM3"   "NCAR-CCSM4"

# using the 5ModelAvg for now 
CR_8.5 <- CR_snapTemp %>%
  filter(scenario == "rcp85" & model == "5ModelAvg")

# reorder by month-date
test <- CR_8.5 %>%
  dplyr::mutate(date = as.Date(date, "%Y%m%d")) %>%
  dplyr::select(date, tasmin, tasmean, tasmax) %>%
  dplyr::arrange(date)

test$std <- (test$tasmax - test$tasmin)/2
test$precision <- (1/(test$std^2))

CR_8.5 <- test
CR_8.5
```

#### Yakutat Forelands

```{r}
YF_snapTemp <- read.csv("DataFiles/YakutatForelands_CreatedFiles/YakutatMonthlyTemperatures.csv", header = TRUE)

# Pull out only the RCP 4.5 data
YF_8.5 <- YF_snapTemp %>%
  filter(scenario == "rcp85")

# Check the model options here
unique(YF_8.5$model) # "5ModelAvg"  "GFDL-CM3"   "NCAR-CCSM4"

# using the 5ModelAvg for now 
YF_8.5 <- YF_snapTemp %>%
  filter(scenario == "rcp85" & model == "5ModelAvg")

# reorder by month-date
test <- YF_8.5 %>%
  dplyr::mutate(date = as.Date(date, "%Y%m%d")) %>%
  dplyr::select(date, tasmin, tasmean, tasmax) %>%
  dplyr::arrange(date)

test$std <- (test$tasmax - test$tasmin)/2
test$precision <- (1/(test$std^2))

YF_8.5 <- test
YF_8.5
```

### Plot the forecasted data

#### Copper River Delta

```{r}
CR_8.5

# Full time frame (2006-2100)
CR_SNAP <- ggplot(data = CR_8.5, mapping = aes(x = date)) +
  geom_line(aes(y = tasmean), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Copper River SNAP Forecast - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
CR_SNAP

ggsave("CR_SNAP_Forecast8.5.jpeg", plot = CR_SNAP, width = 8, height = 6)

# Limited time frame (2020-2100)
test <- CR_8.5 %>%
  filter(date > "2019-12-01")

CR_8.5 <- test

CR_SNAP <- ggplot(data = CR_8.5, mapping = aes(x = date)) +
  geom_line(aes(y = tasmean), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Copper River SNAP Forecast - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
CR_SNAP

ggsave("CR_SNAP_Forecast8.5_2020.jpeg", plot = CR_SNAP, width = 8, height = 6)
```

#### Yakutat Forelands

```{r}
YF_8.5

# Full time frame (2006-2100)
YF_SNAP <- ggplot(data = YF_8.5, mapping = aes(x = date)) +
  geom_line(aes(y = tasmean), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Yakutat Forelands SNAP Forecast - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
YF_SNAP

ggsave("YF_SNAP_Forecast8.5.jpeg", plot = YF_SNAP, width = 8, height = 6)

# Limited time frame (2020-2100)
test <- YF_8.5 %>%
  filter(date > "2019-12-01")

YF_8.5 <- test

YF_SNAP <- ggplot(data = YF_8.5, mapping = aes(x = date)) +
  geom_line(aes(y = tasmean), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Yakutat Forelands SNAP Forecast - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
YF_SNAP

ggsave("YF_SNAP_Forecast8.5_2020.jpeg", plot = YF_SNAP, width = 8, height = 6)
```

# Appending the Temperature Forecasts to the observed data

## Starting with RCP 4.5

### Creating a dataframe with missing values for all other aspects

Determining the columns needed for this

```{r}
CR_ObsData <- read.csv("DataFiles/CopperRiver_CreatedFiles/CR_PondAirTemp_mean.csv", header=TRUE)
head(CR_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CR_ObsData <- CR_ObsData %>%
  dplyr::rename(date=Date)
CR_ObsData <- CR_ObsData[,-1]
head(CR_ObsData)

CR_ObsData <- CR_ObsData %>%
  dplyr::select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CR_ObsData
```

Columns needed are date, Airport, Air_MonthAvg, Air_MonthMax, Air_MonthMin, Region, TIN, SQR, TIS, CAB, EYS, RHM, BVS, CAN, WDD

#### Copper River Delta

Checking the dimensions of the dataframe

```{r}
CR_4.5
dim(CR_4.5) # 972 rows here, so need to make 972 blank rows
```

Checking the observational data

```{r}
CR_ObsData <- read.csv("DataFiles/CopperRiver_CreatedFiles/CR_PondAirTemp_mean.csv", header=TRUE)
head(CR_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CR_ObsData <- CR_ObsData %>%
  dplyr::rename(date=Date)
CR_ObsData <- CR_ObsData[,-1]
head(CR_ObsData)

CR_ObsData <- CR_ObsData %>%
  dplyr::select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CR_ObsData
```

Creating an empty dataframe

```{r}
CR_df <- CR_ObsData[0,]
str(CR_df)

attr(CR_df, 'row.names') <- 1:nrow(CR_4.5)

str(CR_df)
```

#### Yakutat Forelands

Checking the dimensions of the dataframe

```{r}
YF_4.5
dim(YF_4.5) # 972 rows here, so need to make 972 blank rows
```

Checking the observational data

```{r}
YF_ObsData <- read.csv("DataFiles/YakutatForelands_CreatedFiles/YF_PondAirTemp_mean.csv", header=TRUE)
head(YF_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
YF_ObsData <- YF_ObsData %>%
  dplyr::rename(date=Date)
YF_ObsData <- YF_ObsData[,-1]
head(YF_ObsData)

YF_ObsData <- YF_ObsData %>%
  dplyr::select(date, Airport, Air_MonthAvg, std, precision, Region,  MP1, MP3, PL3, PL2, UBP3, MP5, MP8, UBP1, UBP2, PL1, UBP4)
YF_ObsData
```

Creating an empty dataframe

```{r}
YF_df <- YF_ObsData[0,]
str(YF_df)

attr(YF_df, 'row.names') <- 1:nrow(YF_4.5)

str(YF_df)
```

### Adding in the data where it is available

#### Copper River Delta

```{r}
CR_df

#values where there is data
CR_df$date <- CR_4.5$date
CR_df$Air_MonthAvg <- CR_4.5$tasmean
CR_df$std <- CR_4.5$std
CR_df$precision <- CR_4.5$precision
CR_df$Region <- rep("CR", nrow(CR_df))

#values where there is not data
CR_df$Airport <- rep(NA, nrow(CR_df))
CR_df$BVS <- rep(NA, nrow(CR_df))
CR_df$CAB <- rep(NA, nrow(CR_df))
CR_df$CAN <- rep(NA, nrow(CR_df))
CR_df$EYS <- rep(NA, nrow(CR_df))
CR_df$RHM <- rep(NA, nrow(CR_df))
CR_df$SQR <- rep(NA, nrow(CR_df))
CR_df$TIN <- rep(NA, nrow(CR_df))
CR_df$TIS <- rep(NA, nrow(CR_df))
CR_df$WDD <- rep(NA, nrow(CR_df))

str(CR_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
CR_df$date
CR_ObsData$date <- as.Date(CR_ObsData$date)

CR_new_df <- rbind(CR_ObsData, CR_df)

str(CR_new_df)

CR_45Forecast <- CR_new_df
CR_45Forecast

#save this as a .csv file
write.csv(CR_45Forecast, file = "CR_45SNAPForecast.csv")

```

#### Yakutat Forelands

Adding in the data where it is available

```{r}


#values where there is data
YF_df$date <- YF_4.5$date
YF_df$Air_MonthAvg <- YF_4.5$tasmean
YF_df$std <- YF_4.5$std
YF_df$precision <- YF_4.5$precision
YF_df$Region <- rep("YF", nrow(YF_df))

#values where there is not data
YF_df$Airport <- rep(NA, nrow(YF_df))
YF_df$MP1<- rep(NA, nrow(YF_df))
YF_df$MP3 <- rep(NA, nrow(YF_df))
YF_df$MP5 <- rep(NA, nrow(YF_df))
YF_df$MP8 <- rep(NA, nrow(YF_df))
YF_df$PL1 <- rep(NA, nrow(YF_df))
YF_df$PL2 <- rep(NA, nrow(YF_df))
YF_df$PL3 <- rep(NA, nrow(YF_df))
YF_df$UBP1 <- rep(NA, nrow(YF_df))
YF_df$UBP2 <- rep(NA, nrow(YF_df))
YF_df$UBP3 <- rep(NA, nrow(YF_df))
YF_df$UBP4 <- rep(NA, nrow(YF_df))

str(YF_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
YF_df
YF_ObsData$date <- as.Date(YF_ObsData$date)

YF_new_df <- rbind(YF_ObsData, YF_df)

str(YF_new_df)

YF_45Forecast <- YF_new_df
YF_45Forecast

#save this as a .csv file
write.csv(YF_45Forecast, file = "YF_45SNAPForecast.csv")

```

## Continuing with RCP 8.5

### Creating a dataframe with missing values for all other aspects

Determining the columns needed for this

```{r}
CR_ObsData <- read.csv("DataFiles/CopperRiver_CreatedFiles/CR_PondAirTemp_mean.csv", header=TRUE)
head(CR_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CR_ObsData <- CR_ObsData %>%
  dplyr::rename(date=Date)
CR_ObsData <- CR_ObsData[,-1]
head(CR_ObsData)

CR_ObsData <- CR_ObsData %>%
  select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CR_ObsData
```

Columns needed are date, Airport, Air_MonthAvg, Air_MonthMax, Air_MonthMin, Region, TIN, SQR, TIS, CAB, EYS, RHM, BVS, CAN, WDD

#### Copper River Delta

Checking the dimensions of the dataframe

```{r}
CR_8.5
dim(CR_8.5) # 972 rows here, so need to make 972 blank rows
```

Checking the observational data

```{r}
CR_ObsData <- read.csv("DataFiles/CopperRiver_CreatedFiles/CR_PondAirTemp_mean.csv", header=TRUE)
head(CR_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CR_ObsData <- CR_ObsData %>%
  dplyr::rename(date=Date)
CR_ObsData <- CR_ObsData[,-1]
head(CR_ObsData)

CR_ObsData <- CR_ObsData %>%
  select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CR_ObsData
```

Creating an empty dataframe

```{r}
CR_df <- CR_ObsData[0,]
str(CR_df)

attr(CR_df, 'row.names') <- 1:nrow(CR_8.5)

str(CR_df)
```

#### Yakutat Forelands

Checking the dimensions of the dataframe

```{r}
YF_8.5
dim(YF_8.5) # 972 rows here, so need to make 972 blank rows
```

Checking the observational data

```{r}
YF_ObsData <- read.csv("DataFiles/YakutatForelands_CreatedFiles/YF_PondAirTemp_mean.csv", header=TRUE)
head(YF_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
YF_ObsData <- YF_ObsData %>%
  dplyr::rename(date=Date)
YF_ObsData <- YF_ObsData[,-1]
head(YF_ObsData)

YF_ObsData <- YF_ObsData %>%
  select(date, Airport, Air_MonthAvg, std, precision, Region,  MP1, MP3, PL3, PL2, UBP3, MP5, MP8, UBP1, UBP2, PL1, UBP4)
YF_ObsData
```

Creating an empty dataframe

```{r}
YF_df <- YF_ObsData[0,]
str(YF_df)

attr(YF_df, 'row.names') <- 1:nrow(YF_8.5)

str(YF_df)
```

### Adding in the data where it is available

#### Copper River Delta

```{r}
CR_df

#values where there is data
CR_df$date <- CR_8.5$date
CR_df$Air_MonthAvg <- CR_8.5$tasmean
CR_df$std <- CR_8.5$std
CR_df$precision <- CR_8.5$precision
CR_df$Region <- rep("CR", nrow(CR_df))

#values where there is not data
CR_df$Airport <- rep(NA, nrow(CR_df))
CR_df$BVS <- rep(NA, nrow(CR_df))
CR_df$CAB <- rep(NA, nrow(CR_df))
CR_df$CAN <- rep(NA, nrow(CR_df))
CR_df$EYS <- rep(NA, nrow(CR_df))
CR_df$RHM <- rep(NA, nrow(CR_df))
CR_df$SQR <- rep(NA, nrow(CR_df))
CR_df$TIN <- rep(NA, nrow(CR_df))
CR_df$TIS <- rep(NA, nrow(CR_df))
CR_df$WDD <- rep(NA, nrow(CR_df))

str(CR_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
CR_df
CR_ObsData$date <- as.Date(CR_ObsData$date)

CR_new_df <- rbind(CR_ObsData, CR_df)

str(CR_new_df)

CR_85Forecast <- CR_new_df
CR_8.5

#save this as a .csv file
write.csv(CR_85Forecast, file = "CR_85SNAPForecast.csv")

```

#### Yakutat Forelands

Adding in the data where it is available

```{r}
YF_df

#values where there is data
YF_df$date <- YF_8.5$date
YF_df$Air_MonthAvg <- YF_8.5$tasmean
YF_df$std <- YF_8.5$std
YF_df$precision <- YF_8.5$precision
YF_df$Region <- rep("YF", nrow(YF_df))

#values where there is not data
YF_df$Airport <- rep(NA, nrow(YF_df))
YF_df$MP1<- rep(NA, nrow(YF_df))
YF_df$MP3 <- rep(NA, nrow(YF_df))
YF_df$MP5 <- rep(NA, nrow(YF_df))
YF_df$MP8 <- rep(NA, nrow(YF_df))
YF_df$PL1 <- rep(NA, nrow(YF_df))
YF_df$PL2 <- rep(NA, nrow(YF_df))
YF_df$PL3 <- rep(NA, nrow(YF_df))
YF_df$UBP1 <- rep(NA, nrow(YF_df))
YF_df$UBP2 <- rep(NA, nrow(YF_df))
YF_df$UBP3 <- rep(NA, nrow(YF_df))
YF_df$UBP4 <- rep(NA, nrow(YF_df))

str(YF_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
YF_df
YF_ObsData$date <- as.Date(YF_ObsData$date)

YF_new_df <- rbind(YF_ObsData, YF_df)

str(YF_new_df)

YF_85Forecast <- YF_new_df
YF_85Forecast

#save this as a .csv file
write.csv(YF_85Forecast, file = "YF_85SNAPForecast.csv")

```

# Attemping a quick plot of this data for both CR and YF

## RCP 4.5

### Copper River Delta

```{r}
CR_45Forecast 

CR_45Temps<- ggplot(data = CR_45Forecast, mapping = aes(x = date)) +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Copper River Delta Temperature - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
CR_45Temps

ggsave("CR_45Temps.jpeg", plot = CR_45Temps, width = 8, height = 6)
```

### Yakutat Forelands

```{r}
YF_45Forecast 

YF_45Temps<- ggplot(data = YF_45Forecast, mapping = aes(x = date)) +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Yakutat Forelands Temperature - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
YF_45Temps

ggsave("YF_45Temps.jpeg", plot = YF_45Temps, width = 8, height = 6)
```

## RCP 8.5

### Copper River Delta

```{r}
CR_85Forecast 

CR_85Temps<- ggplot(data = CR_85Forecast, mapping = aes(x = date)) +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Copper River Delta Temperature - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
CR_85Temps

ggsave("CR_85Temps.jpeg", plot = CR_85Temps, width = 8, height = 6)
```

### Yakutat Forelands

```{r}
YF_85Forecast 

YF_85Temps<- ggplot(data = YF_85Forecast, mapping = aes(x = date)) +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Yakutat Forelands Temperature - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
YF_85Temps

ggsave("YF_85Temps.jpeg", plot = YF_85Temps, width = 8, height = 6)
```
