---
title: "R Notebook"
output: html_notebook
---

This file is to take the code from DeltaCals_Seasonality_CRD/YF85, DeltaCalcs_Seasonality_CRD/YF45 to make a panel plot for the seasonal change and decadal changes

### Starting with the CRD

```{r}

rm(list = ls())
```

This code is *directly from the code files listed above*
- .csv files were made at the end of the "DeltaCalcs_Seasonality" files
- those are loaded and used here
  
# Reading in the data files

```{r}
#2030
MonthlyStatsCRD45_2030 <- read.csv("MonthlyStatsCRD45_2030.csv") 
MonthlyStatsCRD85_2030 <- read.csv("MonthlyStatsCRD85_2030.csv") 

#2060
MonthlyStatsCRD45_2060 <- read.csv("MonthlyStatsCRD45_2060.csv") 
MonthlyStatsCRD85_2060 <- read.csv("MonthlyStatsCRD85_2060.csv") 

#2090
MonthlyStatsCRD45_2090 <- read.csv("MonthlyStatsCRD45_2090.csv") 
MonthlyStatsCRD85_2090 <- read.csv("MonthlyStatsCRD85_2090.csv") 

```

# Creating the plots for 8.5
2030s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsCRD85_2030, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsCRD85_2030$Month <- as.numeric(MonthlyStatsCRD85_2030$Month)

# Plot using ggplot
MonthlyStatsCRD85_2030Plot <- ggplot(data = MonthlyStatsCRD85_2030, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
MonthlyStatsCRD85_2030Plot
```

2060s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsCRD85_2060, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsCRD85_2060$Month <- as.numeric(MonthlyStatsCRD85_2060$Month)

# Plot using ggplot
MonthlyStatsCRD85_2060Plot <- ggplot(data = MonthlyStatsCRD85_2060, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStatsCRD85_2060Plot
```

2090s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsCRD85_2090, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsCRD85_2090$Month <- as.numeric(MonthlyStatsCRD85_2090$Month)

# Plot using ggplot
MonthlyStatsCRD85_2090Plot <- ggplot(data = MonthlyStatsCRD85_2090, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStatsCRD85_2090Plot
```

Panel Plot
```{r}
ggarrange(MonthlyStatsCRD85_2030Plot, MonthlyStatsCRD85_2060Plot, MonthlyStatsCRD85_2090Plot, ncol = 3)
```

# Creating the plots for 4.5

2030s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsCRD45_2030, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsCRD45_2030$Month <- as.numeric(MonthlyStatsCRD45_2030$Month)

# Plot using ggplot
MonthlyStatsCRD45_2030Plot <- ggplot(data = MonthlyStatsCRD45_2030, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
MonthlyStatsCRD45_2030Plot
```

2060s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsCRD45_2060, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsCRD45_2060$Month <- as.numeric(MonthlyStatsCRD45_2060$Month)

# Plot using ggplot
MonthlyStatsCRD45_2060Plot <- ggplot(data = MonthlyStatsCRD45_2060, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStatsCRD45_2060Plot
```

2090s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsCRD45_2090, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsCRD45_2090$Month <- as.numeric(MonthlyStatsCRD45_2090$Month)

# Plot using ggplot
MonthlyStatsCRD45_2090Plot <- ggplot(data = MonthlyStatsCRD45_2090, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStatsCRD45_2090Plot
```

Panel Plot
```{r}
ggarrange(MonthlyStatsCRD45_2030Plot, MonthlyStatsCRD45_2060Plot, MonthlyStatsCRD45_2090Plot, ncol = 3)
```

# Creating plots with both RCP 4.5 and 8.5

2030s
```{r}
MonthlyStatsCRD_2030Plot <- ggplot() +
  geom_line(data = MonthlyStatsCRD45_2030, aes(x = Month, y = mean_diff_AvgTemp, color = "4.5")) +
  geom_line(data = MonthlyStatsCRD85_2030, aes(x = Month, y = mean_diff_AvgTemp, color = "8.5")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  geom_ribbon(data = MonthlyStatsCRD45_2030, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +
  geom_ribbon(data = MonthlyStatsCRD85_2030, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  ylim(-3, 6) +
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)", color = "RCP") +  # Setting legend title
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  scale_color_manual(values = c("4.5" = "blue", "8.5" = "red"), labels = c("4.5", "8.5")) +  # Adjusting color scale
  theme(legend.position = "none")  # Removing the legend

MonthlyStatsCRD_2030Plot
```

2060s
```{r}
MonthlyStatsCRD_2060Plot <- ggplot() +
  geom_line(data = MonthlyStatsCRD45_2060, aes(x = Month, y = mean_diff_AvgTemp, color = "4.5")) +
  geom_line(data = MonthlyStatsCRD85_2060, aes(x = Month, y = mean_diff_AvgTemp, color = "8.5")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  geom_ribbon(data = MonthlyStatsCRD45_2060, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +
  geom_ribbon(data = MonthlyStatsCRD85_2060, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  ylim(-3, 6) +
  theme_classic() +
  labs(x = "", y = "", color = "RCP") + # Setting legend title
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank()) +
  scale_color_manual(values = c("4.5" = "blue", "8.5" = "red"), labels = c("4.5", "8.5")) + # Adjusting color scale  
  theme(legend.position = "none")  # Removing the legend

MonthlyStatsCRD_2060Plot
```

2090s
```{r}
MonthlyStatsCRD_2090Plot <- ggplot() +
  geom_line(data = MonthlyStatsCRD45_2090, aes(x = Month, y = mean_diff_AvgTemp, color = "4.5")) +
  geom_line(data = MonthlyStatsCRD85_2090, aes(x = Month, y = mean_diff_AvgTemp, color = "8.5")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  geom_ribbon(data = MonthlyStatsCRD45_2090, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +
  geom_ribbon(data = MonthlyStatsCRD85_2090, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  ylim(-3, 6) +
  theme_classic() +
  labs(x = "", y = "", color = "RCP") + #setting the legend title
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())+
  scale_color_manual(values = c("4.5" = "blue", "8.5" = "red"), labels = c("4.5", "8.5")) + # Adjusting color scale  
  theme(legend.position = "none")  # Removing the legend

MonthlyStatsCRD_2090Plot
```

Panel Plot
```{r}
MonthlyStatsCRD <- ggarrange(MonthlyStatsCRD_2030Plot, MonthlyStatsCRD_2060Plot, MonthlyStatsCRD_2090Plot, ncol = 3)
MonthlyStatsCRD
ggsave("DeltaMonthCRD.png", plot = MonthlyStatsCRD, width = 12, height = 3)
```

### Moving on to YF

This code is *directly from the code files listed above*
- .csv files were made at the end of the "DeltaCalcs_Seasonality" files
- those are loaded and used here
  
# Reading in the data files

```{r}
#2030
MonthlyStatsYF45_2030 <- read.csv("MonthlyStatsYF45_2030.csv") 
MonthlyStatsYF85_2030 <- read.csv("MonthlyStatsYF85_2030.csv") 

#2060
MonthlyStatsYF45_2060 <- read.csv("MonthlyStatsYF45_2060.csv") 
MonthlyStatsYF85_2060 <- read.csv("MonthlyStatsYF85_2060.csv") 

#2090
MonthlyStatsYF45_2090 <- read.csv("MonthlyStatsYF45_2090.csv") 
MonthlyStatsYF85_2090 <- read.csv("MonthlyStatsYF85_2090.csv") 

```

# Creating the plots for 8.5
2030s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsYF85_2030, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsYF85_2030$Month <- as.numeric(MonthlyStatsYF85_2030$Month)

# Plot using ggplot
MonthlyStatsYF85_2030Plot <- ggplot(data = MonthlyStatsYF85_2030, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
MonthlyStatsYF85_2030Plot
```

2060s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsYF85_2060, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsYF85_2060$Month <- as.numeric(MonthlyStatsYF85_2060$Month)

# Plot using ggplot
MonthlyStatsYF85_2060Plot <- ggplot(data = MonthlyStatsYF85_2060, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStatsYF85_2060Plot
```

2090s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsYF85_2090, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsYF85_2090$Month <- as.numeric(MonthlyStatsYF85_2090$Month)

# Plot using ggplot
MonthlyStatsYF85_2090Plot <- ggplot(data = MonthlyStatsYF85_2090, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStatsYF85_2090Plot
```

Panel Plot
```{r}
ggarrange(MonthlyStatsYF85_2030Plot, MonthlyStatsYF85_2060Plot, MonthlyStatsYF85_2090Plot, ncol = 3)
```

# Creating the plots for 4.5

2030s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsYF45_2030, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsYF45_2030$Month <- as.numeric(MonthlyStatsYF45_2030$Month)

# Plot using ggplot
MonthlyStatsYF45_2030Plot <- ggplot(data = MonthlyStatsYF45_2030, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
MonthlyStatsYF45_2030Plot
```

2060s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsYF45_2060, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsYF45_2060$Month <- as.numeric(MonthlyStatsYF45_2060$Month)

# Plot using ggplot
MonthlyStatsYF45_2060Plot <- ggplot(data = MonthlyStatsYF45_2060, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStatsYF45_2060Plot
```

2090s
```{r}
# 2. Monthly averages across ponds
ggplot(data = MonthlyStatsYF45_2090, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStatsYF45_2090$Month <- as.numeric(MonthlyStatsYF45_2090$Month)

# Plot using ggplot
MonthlyStatsYF45_2090Plot <- ggplot(data = MonthlyStatsYF45_2090, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-3,6)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStatsYF45_2090Plot
```

Panel Plot
```{r}
ggarrange(MonthlyStatsYF45_2030Plot, MonthlyStatsYF45_2060Plot, MonthlyStatsYF45_2090Plot, ncol = 3)
```

# Creating plots with both RCP 4.5 and 8.5

2030s
```{r}
MonthlyStatsYF_2030Plot <- ggplot() +
  geom_line(data = MonthlyStatsYF45_2030, aes(x = Month, y = mean_diff_AvgTemp, color = "4.5")) +
  geom_line(data = MonthlyStatsYF85_2030, aes(x = Month, y = mean_diff_AvgTemp, color = "8.5")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  geom_ribbon(data = MonthlyStatsYF45_2030, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +
  geom_ribbon(data = MonthlyStatsYF85_2030, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  ylim(-3, 6) +
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)", color = "RCP") +  # Setting legend title
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  scale_color_manual(values = c("4.5" = "blue", "8.5" = "red"), labels = c("4.5", "8.5")) +  # Adjusting color scale
  theme(legend.position = "none")  # Removing the legend

MonthlyStatsYF_2030Plot
```

2060s
```{r}
MonthlyStatsYF_2060Plot <- ggplot() +
  geom_line(data = MonthlyStatsYF45_2060, aes(x = Month, y = mean_diff_AvgTemp, color = "4.5")) +
  geom_line(data = MonthlyStatsYF85_2060, aes(x = Month, y = mean_diff_AvgTemp, color = "8.5")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  geom_ribbon(data = MonthlyStatsYF45_2060, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +
  geom_ribbon(data = MonthlyStatsYF85_2060, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  ylim(-3, 6) +
  theme_classic() +
  labs(x = "", y = "", color = "RCP") + # Setting legend title
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank()) +
  scale_color_manual(values = c("4.5" = "blue", "8.5" = "red"), labels = c("4.5", "8.5")) + # Adjusting color scale  
  theme(legend.position = "none")  # Removing the legend

MonthlyStatsYF_2060Plot
```

2090s
```{r}
MonthlyStatsYF_2090Plot <- ggplot() +
  geom_line(data = MonthlyStatsYF45_2090, aes(x = Month, y = mean_diff_AvgTemp, color = "4.5")) +
  geom_line(data = MonthlyStatsYF85_2090, aes(x = Month, y = mean_diff_AvgTemp, color = "8.5")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  geom_ribbon(data = MonthlyStatsYF45_2090, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +
  geom_ribbon(data = MonthlyStatsYF85_2090, aes(x = Month, ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "red", alpha = 0.3) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  ylim(-3, 6) +
  theme_classic() +
  labs(x = "", y = "", color = "RCP") + #setting the legend title
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())+
  scale_color_manual(values = c("4.5" = "blue", "8.5" = "red"), labels = c("4.5", "8.5")) + # Adjusting color scale  
  theme(legend.position = "none")  # Removing the legend

MonthlyStatsYF_2090Plot
```

Panel Plot
```{r}
MonthlyStatsYF <- ggarrange(MonthlyStatsYF_2030Plot, MonthlyStatsYF_2060Plot, MonthlyStatsYF_2090Plot, ncol = 3)
MonthlyStatsYF
ggsave("DeltaMonthYF.png", plot = MonthlyStatsYF, width = 12, height = 3)
```

