---
title: "Pond Temps - Delta Graphs and Seasonal Trends"
output: html_notebook
---
This is code used to make plots based on output from the "IndividualPondModels_Forecasts45/85" for boththe CR and YF
Read in the output files for each pond/RCP combination then work on delta figures and figures like Hamlet et al. 2020

```{r}
library(dplyr)

```

Starting with the Copper River Delta

Starting with RCP 45

# Reading in the data
```{r}
BVSMod <- read.csv("BVSMod_45.csv", header=TRUE)
BVSMod$date <- as.Date(BVSMod$date)

CABMod <- read.csv("CABMod_45.csv", header=TRUE)
CABMod$date <- as.Date(CABMod$date)

CANMod <- read.csv("CANMod_45.csv", header=TRUE)
CANMod$date <- as.Date(CANMod$date)

EYSMod <- read.csv("EYSMod_45.csv", header=TRUE)
EYSMod$date <- as.Date(EYSMod$date)

RHMMod <- read.csv("RHMMod_45.csv", header=TRUE)
RHMMod$date <- as.Date(RHMMod$date)

SQRMod <- read.csv("SQRMod_45.csv", header=TRUE)
SQRMod$date <- as.Date(SQRMod$date)

TINMod <- read.csv("TINMod_45.csv", header=TRUE)
TINMod$date <- as.Date(TINMod$date)

TISMod <- read.csv("TISMod_45.csv", header=TRUE)
TISMod$date <- as.Date(TISMod$date)

WDDMod <- read.csv("WDDMod_45.csv", header=TRUE)
WDDMod$date <- as.Date(WDDMod$date)
```

 
### 2012-2020
```{r}
# List of pond dataset names
pond_names <- c("BVSMod", "CABMod", "CANMod", "EYSMod", "RHMMod", "SQRMod", "TINMod", "TISMod", "WDDMod")

# Initialize an empty dataframe to store results
avgX50_before_2021 <- data.frame(Pond = pond_names, Avg_X50_before_2021 = numeric(length(pond_names)))

# Loop through each pond dataset
for (i in seq_along(pond_names)) {
  pond_name <- pond_names[i]
  
  # Read the CSV file and convert 'date' column to Date format
  assign(pond_name, read.csv(paste0(pond_name, "_45.csv"), header=TRUE))
  assign(paste0(pond_name, "$date"), as.Date(get(pond_name)$date))
  
  # Filter data for dates before January 1, 2021
  filtered_data <- get(pond_name) %>%
    filter(date < as.Date("2021-01-01"))
  
  # Calculate the mean of X50.
  avg_X50 <- mean(filtered_data$X50., na.rm = TRUE)
  
  # Store the result in the dataframe
  avgX50_before_2021[i, "Avg_X50_before_2021"] <- avg_X50
}

# Print the dataframe
print(avgX50_before_2021)
```

### 2030-2039
```{r}
# Initialize a list to store results
results2030_2039 <- list()

# Loop through each year from 2030 to 2039
for (yr in 2030:2039) {
  yearly_results <- list()
  
  # Loop through each pond dataset
  for (pond_name in c("BVSMod", "CABMod", "CANMod", "EYSMod", "RHMMod", "SQRMod", "TINMod", "TISMod", "WDDMod")) {
    pond_data <- get(pond_name)  # Fetch the dataset by name
    
    # Filter data for the current year
    yearly_data <- pond_data %>%
      filter(year(date) == yr)
    
    # Calculate the mean of X50. for the current year
    yearly_mean <- mean(yearly_data$X50., na.rm = TRUE)
    
    # Store the result in the list
    yearly_results[[pond_name]] <- yearly_mean
  }
  
  # Store yearly results for the current year
  results2030_2039[[as.character(yr)]] <- yearly_results
}
```

### 2060-2069
```{r}
# Initialize a list to store results
results2060_2069 <- list()

# Loop through each year from 2060 to 2069
for (yr in 2060:2069) {
  yearly_results <- list()
  
  # Loop through each pond dataset
  for (pond_name in c("BVSMod", "CABMod", "CANMod", "EYSMod", "RHMMod", "SQRMod", "TINMod", "TISMod", "WDDMod")) {
    pond_data <- get(pond_name)  # Fetch the dataset by name
    
    # Filter data for the current year
    yearly_data <- pond_data %>%
      filter(year(date) == yr)
    
    # Calculate the mean of X50. for the current year
    yearly_mean <- mean(yearly_data$X50., na.rm = TRUE)
    
    # Store the result in the list
    yearly_results[[pond_name]] <- yearly_mean
  }
  
  # Store yearly results for the current year
  results2060_2069[[as.character(yr)]] <- yearly_results
}
```

### 2090-2099
```{r}
# Initialize a list to store results
results2090_2099 <- list()

# Loop through each year from 2090 to 2099
for (yr in 2090:2099) {
  yearly_results <- list()
  
  # Loop through each pond dataset
  for (pond_name in c("BVSMod", "CABMod", "CANMod", "EYSMod", "RHMMod", "SQRMod", "TINMod", "TISMod", "WDDMod")) {
    pond_data <- get(pond_name)  # Fetch the dataset by name
    
    # Filter data for the current year
    yearly_data <- pond_data %>%
      filter(year(date) == yr)
    
    # Calculate the mean of X50. for the current year
    yearly_mean <- mean(yearly_data$X50., na.rm = TRUE)
    
    # Store the result in the list
    yearly_results[[pond_name]] <- yearly_mean
  }
  
  # Store yearly results for the current year
  results2090_2099[[as.character(yr)]] <- yearly_results
}
```

# Put these values for 2030, 2060, and 2090 into the dataframe
```{r}
# List of pond dataset names
pond_names <- c("BVSMod", "CABMod", "CANMod", "EYSMod", "RHMMod", "SQRMod", "TINMod", "TISMod", "WDDMod")

# Initialize an empty dataframe to store results
avgX50_table <- data.frame(Pond = pond_names)

# Create columns for each year from 2030 to 2039, 2060 to 2069, and 2090 to 2099
for (yr in c(2030:2039, 2060:2069, 2090:2099)) {
  col_name <- paste0("Y", yr)  # Column name like "Y2030", "Y2031", etc.
  avgX50_table[[col_name]] <- numeric(length(pond_names))  # Initialize column with zeros
}

# Loop through each pond dataset
for (i in seq_along(pond_names)) {
  pond_name <- pond_names[i]
  
  # Read the CSV file and convert 'date' column to Date format
  assign(pond_name, read.csv(paste0(pond_name, "_45.csv"), header=TRUE))
  assign(paste0(pond_name, "$date"), as.Date(get(pond_name)$date))
  
  # Loop through each year in the 2030s, 2060s, and 2090s
  for (yr in c(2030:2039, 2060:2069, 2090:2099)) {
    # Calculate mean X50 values for the current year and store in dataframe
    avgX50_table[i, paste0("Y", yr)] <- mean(get(pond_name) %>%
                                              filter(year(date) == yr) %>%
                                              pull(X50.), na.rm = TRUE)
  }
}

# Print the final table
print(avgX50_table)

```


```{r}
# List of pond dataset names
pond_names <- c("BVS", "CAB", "CAN", "EYS", "RHM", "SQR", "TIN", "TIS", "WDD")

# Initialize an empty dataframe to store results
avgX50_table <- data.frame(Pond = pond_names)

# Add column for Avg_X50_before_2021 as 2020
avgX50_table$`2020` <- numeric(length(pond_names))

# Create columns for each year from 2030 to 2039, 2060 to 2069, and 2090 to 2099
years <- c(2030:2039, 2060:2069, 2090:2099)
for (yr in years) {
  col_name <- paste0("Y", yr)  # Column name like "Y2030", "Y2060", etc.
  avgX50_table[[col_name]] <- numeric(length(pond_names))  # Initialize column with zeros
}

# Loop through each pond dataset
for (i in seq_along(pond_names)) {
  pond_name <- pond_names[i]
  
  # Read the CSV file and convert 'date' column to Date format
  assign(pond_name, read.csv(paste0(pond_name, "Mod_45.csv"), header=TRUE))
  assign(paste0(pond_name, "$date"), as.Date(get(pond_name)$date))
  
  # Calculate mean X50 values before 2021
  avgX50_before_2021 <- mean(get(pond_name) %>%
                              filter(date < as.Date("2021-01-01")) %>%
                              pull(X50.), na.rm = TRUE)
  
  # Store mean X50 value before 2021 in avgX50_table as 2020
  avgX50_table$`2020`[i] <- avgX50_before_2021
  
  # Loop through each year in the 2030s, 2060s, and 2090s
  for (yr in years) {
    # Calculate mean X50 values for the current year and store in dataframe
    avgX50_table[i, paste0("Y", yr)] <- mean(get(pond_name) %>%
                                              filter(year(date) == yr) %>%
                                              pull(X50.), na.rm = TRUE)
  }
}

# Print the final table
print(avgX50_table)
```

## Remove the Y from each column name
```{r}
library(dplyr)

# Function to remove 'Y' from column names
remove_Y_prefix <- function(name) {
  gsub("^Y", "", name)
}

# Apply the function to all column names
avgX50_table <- avgX50_table %>%
  rename_with(remove_Y_prefix, starts_with("Y"))

# Print the structure of the updated dataframe
str(avgX50_table)
```


# Calcuate the differences here as new columns in the dataframe
```{r}
# Assuming avgX50_table is loaded and structure verified
Delta_CRD45 <- avgX50_table  # Use your actual dataframe

# Columns to calculate differences for
years <- c("2030", "2031", "2032", "2033", "2034", 
           "2035", "2036", "2037", "2038", "2039", 
           "2060", "2061", "2062", "2063", "2064", 
           "2065", "2066", "2067", "2068", "2069", 
           "2090", "2091", "2092", "2093", "2094", 
           "2095", "2096", "2097", "2098", "2099")

# Loop through each year and calculate the difference
for (year in years) {
  # Create a new column for the difference
  diff_col_name <- paste0("diff_", year)
  Delta_CRD45[[diff_col_name]] <- Delta_CRD45[[year]] - Delta_CRD45$`2020`
}

# Display the structure of Delta_CRD45 to verify
str(Delta_CRD45)
```

#Calculate the mean differences

```{r}
library(dplyr)
library(matrixStats)  # for rowMeans and rowSds functions

# Selecting columns for the specified year ranges and calculating mean and standard deviation
CRD45_selected <- Delta_CRD45 %>%
  mutate(
    Mean_2030_2039 = rowMeans(select(., starts_with("diff_"), matches("203[0-9]$"))),
    SD_2030_2039 = rowSds(as.matrix(select(., starts_with("diff_"), matches("203[0-9]$")))),
    Mean_2060_2069 = rowMeans(select(., starts_with("diff_"), matches("206[0-9]$"))),
    SD_2060_2069 = rowSds(as.matrix(select(., starts_with("diff_"), matches("206[0-9]$")))),
    Mean_2090_2099 = rowMeans(select(., starts_with("diff_"), matches("209[0-9]$"))),
    SD_2090_2099 = rowSds(as.matrix(select(., starts_with("diff_"), matches("209[0-9]$"))))
  )

# Print the structure of the updated dataframe
str(CRD45_selected)
```

## Put these into total mean/standard deviations

```{r}
# Load necessary libraries
library(dplyr)

# Select and rename columns
CRD45_decadal <- CRD45_selected %>%
  select(Pond, starts_with("Mean_"), starts_with("SD_"))

# Rename columns without the 'Y' prefix for means
names(CRD45_decadal) <- sub("^Mean", "Mean", names(CRD45_decadal))

# Rename columns without the 'Y' prefix for standard deviations
names(CRD45_decadal) <- sub("^SD_", "SD_", names(CRD45_decadal))

# Print the structure of the updated dataframe
str(CRD45_decadal)
```

# Making the plots of these deltas
```{r}
# Load required packages
library(ggplot2)
library(gridExtra)

# Create a grouped bar plot with error bars for Mean 2030
plot_2030 <- ggplot(CRD45_decadal, aes(x = Pond, y = Mean_2030_2039)) +
  geom_bar(stat = "identity", fill = "#FFD92F", color = "black") +
  geom_errorbar(aes(ymin = Mean_2030_2039 - SD_2030_2039, ymax = Mean_2030_2039 + SD_2030_2039),
                width = 0.4,  # Adjust the width of error bars as needed
                position = position_dodge(width = 0.9)) +
  labs(x = "", y = "Delta Temp (C)", title = "2030") +
  ylim(-1, 10) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))  # Centered title
plot_2030

# Create a grouped bar plot with error bars for Mean 2060
plot_2060 <- ggplot(CRD45_decadal, aes(x = Pond, y = Mean_2060_2069)) +
  geom_bar(stat = "identity", fill = "#F46D43", color = "black") +
  geom_errorbar(aes(ymin = Mean_2060_2069 - SD_2060_2069, ymax = Mean_2060_2069 + SD_2060_2069),
                width = 0.4,  # Adjust the width of error bars as needed
                position = position_dodge(width = 0.9)) +
  labs(x = "", y = "", title = "2060") +  # No y-axis label
  ylim(-1, 10) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) +  # Centered title 
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())   # Remove Y-axis ticks
plot_2060

# Create a grouped bar plot with error bars for Mean 2090
plot_2090 <- ggplot(CRD45_decadal, aes(x = Pond, y = Mean_2090_2099)) +
  geom_bar(stat = "identity", fill = "#D73027", color = "black") +
  geom_errorbar(aes(ymin = Mean_2090_2099 - SD_2090_2099, ymax = Mean_2090_2099 + SD_2090_2099),
                width = 0.4,  # Adjust the width of error bars as needed
                position = position_dodge(width = 0.9)) +
  labs(x = "", y = "", title = "2090") +  # No y-axis label
  ylim(-1, 10) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) +  # Centered title 
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())   # Remove Y-axis ticks
plot_2090

# Combine plots into a panel plot
Delta_CRD45_plot <- ggarrange(plot_2030, plot_2060, plot_2090, ncol = 3)

# Save the combined plot
ggsave("Output/GeoSNAP/DeltaCRD45.png", plot = Delta_CRD45_plot, width = 8, height = 3)
```

# Statistical tests for each of these ponds for 2030s, 2060s, and 2090s

### 2030s
```{r}
CRD45_selected

# Load required libraries
library(dplyr)
library(broom)

# Step 1: Prepare the data
data_long <- Delta_CRD45 %>%
  select(Pond, matches("^diff_203[0-9]$")) %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "Difference")

boxplot(data_long$Difference ~ data_long$Pond)

# Step 2: Conduct ANOVA by grouping by Pond
res_CRD45 <- aov(Difference ~ Pond, data = data_long)
summary(res_CRD45)

# Step 3: Tukey HSD
post_test_CRD45 <- TukeyHSD(res_CRD45)
post_test_CRD45

```

### 2060s
```{r}
CRD45_selected

# Load required libraries
library(dplyr)
library(broom)

# Step 1: Prepare the data
data_long <- Delta_CRD45 %>%
  select(Pond, matches("^diff_206[0-9]$")) %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "Difference")

boxplot(data_long$Difference ~ data_long$Pond)

# Step 2: Conduct ANOVA by grouping by Pond
res_CRD45 <- aov(Difference ~ Pond, data = data_long)
summary(res_CRD45)

# Step 3: Tukey HSD
post_test_CRD45 <- TukeyHSD(res_CRD45)
post_test_CRD45
```

### 2090s
```{r}
CRD45_selected

# Load required libraries
library(dplyr)
library(broom)

# Step 1: Prepare the data
data_long <- Delta_CRD45 %>%
  select(Pond, matches("^diff_209[0-9]$")) %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "Difference")

boxplot(data_long$Difference ~ data_long$Pond)

# Step 2: Conduct ANOVA by grouping by Pond
res_CRD45 <- aov(Difference ~ Pond, data = data_long)
summary(res_CRD45)

# Step 3: Tukey HSD
post_test_CRD45 <- TukeyHSD(res_CRD45)
post_test_CRD45
```

# Making plots of error by season

## 2012-2020
```{r}
# List of pond dataset names
pond_names <- c("BVSMod", "CABMod", "CANMod", "EYSMod", "RHMMod", "SQRMod", "TINMod", "TISMod", "WDDMod")

# Short names corresponding to pond_names
short_names <- c("BVS", "CAB", "CAN", "EYS", "RHM", "SQR", "TIN", "TIS", "WDD")

# Initialize an empty dataframe to store monthly mean results
monthly_means <- data.frame(Pond = character(), Month = character(), AvgTemp = numeric(), stringsAsFactors = FALSE)

# Loop through each pond dataset
for (i in seq_along(pond_names)) {
  pond_name <- pond_names[i]
  short_name <- short_names[i]  # Get corresponding short name
  
  # Read the CSV file
  pond_data <- read.csv(paste0(pond_name, "_45.csv"), header=TRUE)
  
  # Convert 'date' column to Date class
  pond_data$date <- as.Date(pond_data$date)
  
  # Loop through each month
  for (month in month.name) {
    # Filter data for the current month
    filtered_data <- subset(pond_data, format(date, "%B") == month | format(date, "%b") == month)
    
    # Calculate the mean of X50 for the current month
    avg_X50 <- mean(filtered_data$X50., na.rm = TRUE)
    
    # Append results to monthly_means dataframe
    monthly_means <- rbind(monthly_means, data.frame(Pond = short_name, Month = month, AvgTemp = avg_X50, Year = 2020))
  }
}

# Print the monthly means dataframe
print(monthly_means)

```
## 2030, 2060, and 2090
```{r}
# List of pond dataset names with "Mod" suffix
pond_names_mod <- c("BVSMod", "CABMod", "CANMod", "EYSMod", "RHMMod", "SQRMod", "TINMod", "TISMod", "WDDMod")

# Desired short names for the ponds
short_names <- c("BVS", "CAB", "CAN", "EYS", "RHM", "SQR", "TIN", "TIS", "WDD")

# Initialize empty dataframes to store results
monthly_avgX50_table_2030s <- data.frame(Pond = character(), Month = character(), AvgTemp = numeric(), Year = integer(), stringsAsFactors = FALSE)
monthly_avgX50_table_2060s <- data.frame(Pond = character(), Month = character(), AvgTemp = numeric(), Year = integer(), stringsAsFactors = FALSE)
monthly_avgX50_table_2090s <- data.frame(Pond = character(), Month = character(), AvgTemp = numeric(), Year = integer(), stringsAsFactors = FALSE)

# Function to calculate monthly averages from yearly results
calculate_monthly_averages <- function(results_list, decade_start, decade_end, pond_names, short_names) {
  # Initialize empty dataframe to store monthly averages
  monthly_avg_table <- data.frame(Pond = character(), Month = character(), AvgTemp = numeric(), Year = integer(), stringsAsFactors = FALSE)
  
  # Loop through each year in the decade
  for (yr in decade_start:decade_end) {
    yearly_results <- results_list[[as.character(yr)]]
    
    # Loop through each pond dataset
    for (i in seq_along(pond_names)) {
      pond_name_mod <- pond_names[i]
      pond_name <- short_names[i]
      
      avg_X50 <- yearly_results[[pond_name_mod]]
      
      # Loop through each month
      for (month in month.name) {
        # Filter data for the current year and month
        monthly_data <- pond_data %>%
          filter(year(date) == yr & format(date, "%B") == month)
        
        # Calculate the mean of X50. for the current month
        if (nrow(monthly_data) > 0) {
          avg_temp <- mean(monthly_data$X50., na.rm = TRUE)
        } else {
          avg_temp <- NA
        }
        
        # Append results to monthly_avg_table dataframe
        monthly_avg_table <- rbind(monthly_avg_table, data.frame(Pond = pond_name, Month = month, AvgTemp = avg_temp, Year = yr))
      }
    }
  }
  
  return(monthly_avg_table)
}

# Calculate monthly averages for each decade
monthly_avgX50_table_2030s <- calculate_monthly_averages(results2030_2039, 2030, 2039, pond_names_mod, short_names)
monthly_avgX50_table_2060s <- calculate_monthly_averages(results2060_2069, 2060, 2069, pond_names_mod, short_names)
monthly_avgX50_table_2090s <- calculate_monthly_averages(results2090_2099, 2090, 2099, pond_names_mod, short_names)

# Print the results for each decade
print("Monthly Average X50 for the 2030s:")
print(monthly_avgX50_table_2030s)

print("Monthly Average X50 for the 2060s:")
print(monthly_avgX50_table_2060s)

print("Monthly Average X50 for the 2090s:")
print(monthly_avgX50_table_2090s)
```
# Combine all the monthly average tables into one dataframe 

```{r}
# Combine all monthly average tables into one dataframe
combined_monthly_avg <- bind_rows(
  monthly_means,
  monthly_avgX50_table_2030s,
  monthly_avgX50_table_2060s,
  monthly_avgX50_table_2090s
)
combined_monthly_avg
```

# Calculate the differences here as compared to 2020
```{r}
# Define columns (months)
months <- unique(combined_monthly_avg$Month)

# Reference year
reference_year <- 2020

# Loop through each month and calculate the difference
for (month in months) {
  # Subset data for the current month
  month_data <- combined_monthly_avg[combined_monthly_avg$Month == month, ]
  
  # Calculate the difference relative to the reference year
  combined_monthly_avg$diff_AvgTemp[combined_monthly_avg$Month == month] <- month_data$AvgTemp - month_data$AvgTemp[month_data$Year == reference_year]
}

# Display the structure of combined_monthly_avg to verify
str(combined_monthly_avg)

DeltaMonthly_CRD45 <- combined_monthly_avg %>%
  filter(Year != 2020)
DeltaMonthly_CRD45
```

# Calculate mean differences

```{r}
library(dplyr)

DeltaMonthly_CRD45

DiffMonthly_CRD45 <- DeltaMonthly_CRD45 %>%
  select(Pond, Month, Year, diff_AvgTemp)

# Create a new column based on year ranges
DiffMonthly_CRD45$Decade <- ifelse(DiffMonthly_CRD45$Year >= 2030 & DiffMonthly_CRD45$Year < 2040, "2030s",
                          ifelse(DiffMonthly_CRD45$Year >= 2060 & DiffMonthly_CRD45$Year < 2070, "2060s",
                          ifelse(DiffMonthly_CRD45$Year >= 2090 & DiffMonthly_CRD45$Year < 2100, "2090s",
                          "Other")))

# Print the updated data frame
print(DiffMonthly_CRD45)
  
# Convert Month to Date type
DiffMonthly_CRD45 <- DiffMonthly_CRD45 %>%
  mutate(Month = as.Date(paste0(Month, " 1 ", Year), format = "%B %d %Y"))

print(DiffMonthly_CRD45)

# Group by Pond and Month, then calculate mean and standard deviation of diff_AvgTemp
summary_stats <- DiffMonthly_CRD45 %>%
  dplyr::mutate(Month = format(Month, "%m")) %>%  # Extracts YYYY-MM format
  dplyr::group_by(Pond, Month, Decade) %>%
  dplyr::summarise(
    mean_diff_AvgTemp = mean(diff_AvgTemp, na.rm = TRUE),
    sd_diff_AvgTemp = sd(diff_AvgTemp, na.rm = TRUE)
  )

# View the summary statistics
print(summary_stats)

# Group by just Month, then calculate mean and standard deviation of diff_AvgTemp
summary_stats2 <- DiffMonthly_CRD45 %>%
  dplyr::mutate(Month = format(Month, "%m")) %>%  # Extracts YYYY-MM format
  dplyr::group_by(Month, Decade) %>%
  dplyr::summarise(
    mean_diff_AvgTemp = mean(diff_AvgTemp, na.rm = TRUE),
    sd_diff_AvgTemp = sd(diff_AvgTemp, na.rm = TRUE)
  )

# View the summary statistics
print(summary_stats2)
```

Plotting the monthly temperature change graphs

2030s
```{r}
summary_stats

# testing out plots

# 1. individual pond with points/0 line
MonthlyStats_2030 <- summary_stats %>%
  filter(Decade == "2030s") 

ggplot(data = MonthlyStats_2030, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp, color = Pond)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStats_2030$Month <- as.numeric(MonthlyStats_2030$Month)

# Plot using ggplot
ggplot(data = MonthlyStats_2030, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp, color = Pond)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)", color = "Pond")+
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))


# 2. Monthly averages across ponds
MonthlyStats_2030 <- summary_stats2 %>%
  filter(Decade == "2030s") 

ggplot(data = MonthlyStats_2030, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStats_2030$Month <- as.numeric(MonthlyStats_2030$Month)

# Plot using ggplot
MonthlyStats_2030Plot <- ggplot(data = MonthlyStats_2030, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "lightblue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))

ggsave("Output/GeoSNAP/MonthlyDeltaCRD45.png", plot = MonthlyStats_2030Plot, width = 8, height = 3)
```
























