---
title: "IndividualPondModels_Observed_YF"
author: "Amaryllis Adey"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goal of this code is to fit a model to the observational data for each of the ponds in the Yakutat Forelands

# Modeling the Observational Data for the Yakutat Forelands Ponds

Looping through each pond individually because the for loop to do this was more complicated than I could figure out in the time before SFS

# Set up the working space for the analysis

## Load in packages

```{r}
library(rjags)
library(daymetr)
library(ecoforecastR)
library(dplyr)
library(lubridate)
library(zoo)
library(ggplot2)
```

## Start by loading in the Yakutat Forelands mean data set

```{r}
YF_Temps <- read.csv("DataFiles/YF_PondAirTemp_mean.csv", header=TRUE)

library(dplyr)
# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
YF_Temps <- YF_Temps %>%
  rename(date=Date)
YF_Temps <- YF_Temps[,-1]
head(YF_Temps)
```

## Separate the data into the air temp and the water temp data

```{r}
YF_AirTemp <- YF_Temps %>%
  select(date, Airport, Air_MonthAvg, Region)

YF_WaterTemp <- YF_Temps %>%
  select(date, MP1, MP3, PL3, PL2, UBP3, MP5, MP8, UBP1, UBP2, PL1, UBP4)
```

## Going to attempt to plot the water temperature and the air temperature over time

```{r}
head(YF_Temps)

# # plotting this --> shows up as 'blank'
# ggplot(YF_Temps)+
#   geom_line(data = YF_Temps, aes(x = date, y = Air_MonthAvg, color = 'red', linetype = "solid"))+
#   geom_line(data = YF_Temps, aes(x = date, y = UBP4, color = 'blue', linetype = "dashed"))

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(tidyr)

# Reshape the data from wide to long format
YF_Temps_long <- YF_Temps %>%
  pivot_longer(cols = c(Air_MonthAvg, MP1, MP3, PL3, PL2, UBP3, MP5, MP8, UBP1, UBP2, PL1, UBP4),
               names_to = "variable",
               values_to = "value")

# Check the structure of the reshaped data
str(YF_Temps_long)

# Create the plot
# Create the plot
YF_PondTemp <- ggplot(YF_Temps_long, aes(x = as.Date(date), y = value, color = variable)) +
  geom_line(data = filter(YF_Temps_long, variable == "Air_MonthAvg"), color = "black") +  
  geom_line(data = filter(YF_Temps_long, variable != "Air_MonthAvg")) +  
  labs(title = "YF Pond Water Temperature Over Time",
       x = "Date",
       y = "Value",
       color = "Variable") +
  theme_minimal()
YF_PondTemp


# Save the plot
ggsave("Plots/YF_PondTemp.png", plot = YF_PondTemp, width = 8, height = 6)

```

# Starting the Analysis

## Part 1 - Pulling out the data that we want
-Linear air-water temperature relationship with only one pond here

# MP1

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- MP1
y <- YF_WaterTemp$MP1
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
MP1ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = MP1)) +
  ggtitle("MP1 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
MP1ModelPlot

MP1ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = MP1)) +
  ggtitle("MP1 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
MP1ModelErrorPlot

ggsave("Plots/MP1ModelPlot.jpeg", plot = MP1ModelPlot, width = 8, height = 6)
ggsave("Plots/MP1ModelErrorPlot.jpeg", plot = MP1ModelErrorPlot, width = 8, height = 6)
```

# MP3

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- MP3
y <- YF_WaterTemp$MP3
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
MP3ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = MP3)) +
  ggtitle("MP3 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
MP3ModelPlot

MP3ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = MP3)) +
  ggtitle("MP3 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
MP3ModelErrorPlot

ggsave("Plots/MP3ModelPlot.jpeg", plot = MP3ModelPlot, width = 8, height = 6)
ggsave("Plots/MP3ModelErrorPlot.jpeg", plot = MP3ModelErrorPlot, width = 8, height = 6)
```

# MP5

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- MP5
y <- YF_WaterTemp$MP5
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
MP5ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = MP5)) +
  ggtitle("MP5 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
MP5ModelPlot

MP5ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = MP5)) +
  ggtitle("MP5 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
MP5ModelErrorPlot

ggsave("Plots/MP5ModelPlot.jpeg", plot = MP5ModelPlot, width = 8, height = 6)
ggsave("Plots/MP5ModelErrorPlot.jpeg", plot = MP5ModelErrorPlot, width = 8, height = 6)
```

# MP8

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- MP8
y <- YF_WaterTemp$MP8
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
MP8ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = MP8)) +
  ggtitle("MP8 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
MP8ModelPlot

MP8ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = MP8)) +
  ggtitle("MP8 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
MP8ModelErrorPlot

ggsave("Plots/MP8ModelPlot.jpeg", plot = MP8ModelPlot, width = 8, height = 6)
ggsave("Plots/MP8ModelErrorPlot.jpeg", plot = MP8ModelErrorPlot, width = 8, height = 6)
```

# PL1

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- PL1
y <- YF_WaterTemp$PL1
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
PL1ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = PL1)) +
  ggtitle("PL1 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
PL1ModelPlot

PL1ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = PL1)) +
  ggtitle("PL1 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
PL1ModelErrorPlot

ggsave("Plots/PL1ModelPlot.jpeg", plot = PL1ModelPlot, width = 8, height = 6)
ggsave("Plots/PL1ModelErrorPlot.jpeg", plot = PL1ModelErrorPlot, width = 8, height = 6)
```

# PL2

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- PL2
y <- YF_WaterTemp$PL2
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
PL2ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = PL2)) +
  ggtitle("PL2 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
PL2ModelPlot

PL2ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = PL2)) +
  ggtitle("PL2 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
PL2ModelErrorPlot

ggsave("Plots/PL2ModelPlot.jpeg", plot = PL2ModelPlot, width = 8, height = 6)
ggsave("Plots/PL2ModelErrorPlot.jpeg", plot = PL2ModelErrorPlot, width = 8, height = 6)
```

# PL3

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- PL3
y <- YF_WaterTemp$PL3
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
PL3ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = PL3)) +
  ggtitle("PL3 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
PL3ModelPlot

PL3ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = PL3)) +
  ggtitle("PL3 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
PL3ModelErrorPlot

ggsave("Plots/PL3ModelPlot.jpeg", plot = PL3ModelPlot, width = 8, height = 6)
ggsave("Plots/PL3ModelErrorPlot.jpeg", plot = PL3ModelErrorPlot, width = 8, height = 6)
```

# UBP1

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- UBP1
y <- YF_WaterTemp$UBP1
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
UBP1ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = UBP1)) +
  ggtitle("UBP1 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
UBP1ModelPlot

UBP1ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = UBP1)) +
  ggtitle("UBP1- Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
UBP1ModelErrorPlot

ggsave("Plots/UBP1ModelPlot.jpeg", plot = UBP1ModelPlot, width = 8, height = 6)
ggsave("Plots/UBP1ModelErrorPlot.jpeg", plot = UBP1ModelErrorPlot, width = 8, height = 6)
```

# UBP2

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- UBP2
y <- YF_WaterTemp$UBP2
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
UBP2ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = UBP2)) +
  ggtitle("UBP2 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
UBP2ModelPlot

UBP2ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = UBP2)) +
  ggtitle("UBP2- Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
UBP2ModelErrorPlot

ggsave("Plots/UBP2ModelPlot.jpeg", plot = UBP2ModelPlot, width = 8, height = 6)
ggsave("Plots/UBP2ModelErrorPlot.jpeg", plot = UBP2ModelErrorPlot, width = 8, height = 6)
```

# UBP3

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- UBP3
y <- YF_WaterTemp$UBP3
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
UBP3ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = UBP3)) +
  ggtitle("UBP3 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
UBP3ModelPlot

UBP3ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = UBP3)) +
  ggtitle("UBP3- Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
UBP3ModelErrorPlot

ggsave("Plots/UBP3ModelPlot.jpeg", plot = UBP3ModelPlot, width = 8, height = 6)
ggsave("Plots/UBP3ModelErrorPlot.jpeg", plot = UBP3ModelErrorPlot, width = 8, height = 6)
```

# UBP4

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# Reset row names
rownames(YF_WaterTemp) <- NULL

# pulling out the dates here
time <- YF_WaterTemp$date
length(time)

# Choose one pond to plot through time -- UBP4
y <- YF_WaterTemp$UBP4
length(y)

# Pull out the air temperature from the data
temp <- YF_AirTemp$Air_MonthAvg
length(temp)
```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk --\> added in the air temperature here

Setup Notes -beta0 as a matrix for the intercept and slope -linear relationship between air and water temperature (k = m\*x+b)

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta0[1] + beta0[2] * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta0 ~ dmnorm(betap, sigmap)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,
             betap=c(1,1), sigmap=diag(0.0001, 2, 2),
             temp = temp, ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta0[1]", "beta0[2]"),
                                n.iter = 1000)
plot(jags.out)
```


### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs", "beta0"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
# THIS CODE DOES NOT WORK -- NEXT CHUNK IS TRYING TO GET THESE SAME PLOTS TO WORK
time                               ## adjust to zoom in and out
time <-as.Date(time)
out <- as.matrix(jags.out)         ## convert from coda to matrix
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols], 2, quantile, c(0.025,0.5,0.975))

# changing ci to a dataframe so that it can be plotted
  # Load necessary libraries
  library(tidyverse)
  
  # Assuming `ci` is your matrix
  ci_df <- as.data.frame(ci)
  names(ci_df) <- time

  # Transpose the dataframe
  ci_transposed <- as.data.frame(t(ci_df))

  # Change the dates to be a column
  ci_transposed <- rownames_to_column(ci_transposed, var = "date")
  ci_transposed$date <- as.Date(ci_transposed$date)
  
  ModelOutput <- ci_transposed
  
  GraphDataOutputCheck <- merge(ModelOutput, YF_WaterTemp, by = "date", all = T)

# Now you can use this data frame with ggplot2
UBP4ModelPlot <- ggplot(GraphDataOutputCheck) +
  geom_line(aes(x = date, y = `50%`)) +
  geom_line(aes(x = date, y = `2.5%`)) +
  geom_line(aes(x = date, y = `97.5%`)) +
  geom_point(aes(x = date, y = UBP4)) +
  ggtitle("UBP4 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
UBP4ModelPlot

UBP4ModelErrorPlot <- ggplot(GraphDataOutputCheck, mapping = aes(date)) +
  geom_line(mapping = aes(y = `2.5%`)) +
  geom_line(mapping = aes(y = `97.5%`)) +
  geom_ribbon(mapping = aes(ymin = `2.5%`, ymax = `97.5%`), fill = "lightgrey") +
  geom_line(aes(x = date, y = `50%`), color = "black") +
  geom_point(aes(x = date, y = UBP4)) +
  ggtitle("UBP4 - Yakutat Forelands")+
  ylab("Temeprature (C)") +
  xlab("Date") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
UBP4ModelErrorPlot

ggsave("Plots/UBP4ModelPlot.jpeg", plot = UBP4ModelPlot, width = 8, height = 6)
ggsave("Plots/UBP4ModelErrorPlot.jpeg", plot = UBP4ModelErrorPlot, width = 8, height = 6)
```