---
title: "StateSpacePondTempModel - YF"
author: "Amaryllis Adey"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exploring the state-space model framework using code based on 'Activity 6 - State-space models' that Alyssa shared along

# Set up the working space for the analysis

## Load in packages

```{r}
library(rjags)
library(daymetr)
library(ecoforecastR)
library(dplyr)
library(lubridate)
library(zoo)
```

## Start by loading in the Yakutat Forelands mean data set

```{r}
YF_Temps <- YF_PondAirTemp_mean
```

## Separate the data into the air temp and the water temp data

```{r}
YF_AirTemp <- YF_Temps %>%
  select(`Month-Year`, Airport, Air_MonthAvg, Region)

YF_WaterTemp <- YF_Temps %>%
  select(`Month-Year`, MP1, MP3, PL3, PL2, UBP3, MP5, MP8, UBP1, UBP2, PL1, UBP4)
```

# Starting the Analysis

## Part 1 - Model with only the water temperature in it

### Pull in the data we want to analyze

Here this is set up as just one pond and temperature data

```{r}
YF_WaterTemp

# pulling out the dates here
time <- as.Date(YF_WaterTemp$`Month-Year`, "%m-%Y")

# Choose one pond to plot through time -- MP1
y = YF_WaterTemp %>%
  select(MP1)

```

The code itself has three components, the data model, the process model, and the priors.

Data Model - relates the observed data, y, at any time point to the latent variable, x. For this example we'll assume that the observation model just consists of Gaussian observation error.

Process Model - relates the state of the system at one point in time to the state one time step ahead. In this case we'll start with the simplest possible process model, a random walk, which just consists of Gaussian process error centered around the current value of the system.

Priors - Finally, for the priors we need to define **priors** for the initial condition, the process error, and the observation error.

### Setting up the random walk

```{r}
RandomWalk = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(x[t-1],tau_add)
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

```{r}
data <- list(y=y$MP1,n=length(y),     ## data
             x_ic=1,tau_ic=1,           ## initial condition prior
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y$MP1)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs"),
                                n.iter = 1000)
plot(jags.out)
```

### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
time.rng = c(1,length(time))       ## adjust to zoom in and out
out <- as.matrix(jags.out)         ## convert from coda to matrix  
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(out[,x.cols],2,quantile,c(0.025,0.5,0.975)) ## model was fit on log scale -- AA think I can skip this since I got rid of the log transformation
ci <- apply(data.frame(out[, x.cols]), 2, quantile, c(0.025, 0.5, 0.975))

plot(time,ci[2,],type='n',ylim=range(y,na.rm=TRUE),ylab="Water Temp",log='y',xlim=time[time.rng])
## adjust x-axis label to be monthly if zoomed
if(diff(time.rng) < 100){ 
  axis.Date(1, at=seq(time[time.rng[1]],time[time.rng[2]],by='month'), format = "%Y-%m")
}
ecoforecastR::ciEnvelope(time,ci[1,],ci[3,],col=ecoforecastR::col.alpha("lightBlue",0.75))
points(time,y,pch="+",cex=0.5)
```

### Next, lets look at the posterior distributions for `tau_add` and `tau_obs`, which we'll convert from precisions back into **standard deviations**.

```{r}
hist(1/sqrt(out[,1]),main=colnames(out)[1])
hist(1/sqrt(out[,2]),main=colnames(out)[2])
```

### We'll also want to look at the joint distribution of the two parameters to check whether the two parameters strongly covary.

```{r, fig.asp = 1.0}
plot(out[,1],out[,2],pch=".",xlab=colnames(out)[1],ylab=colnames(out)[2])
cor(out[,1:2])
```

## Part 2 - Adding in more than one pond

```{r}
YF_WaterTemp

# pulling out the dates here
time <- as.Date(YF_WaterTemp$`Month-Year`, "%m-%Y")

# Choose one pond to plot through time -- MP1
y = YF_WaterTemp %>%
  select(MP1, MP3)
```

### Trying to include more of the ponds here instead of just the one like above

```{r}
RandomWalk_all = "
model{
  
  #### Data Model
  for(p in 1:q){
   for(t in 1:n){
      y[t,p] ~ dnorm(x[t,p],tau_obs)
  }
  }
  
  
  #### Process Model
  for(p in 1:q){
   for(t in 2:n){
    x[t,p]~dnorm(x[t-1,p],tau_add)
  }
  }
 
  
  #### Priors
  for(p in 1:q){
  x[1,p] ~ dnorm(x_ic,tau_ic)
  }
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

This should be for the RandonWalk_all now

```{r}
data <- list(y=y, n=nrow(y), q=ncol(y),    ## data
             x_ic=0,tau_ic=0.2,
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y$MP1, y$MP3)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_all),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs"),
                                n.iter = 1000)
plot(jags.out)
```

### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
time.rng = c(1,length(time))       ## adjust to zoom in and out
out <- as.matrix(jags.out)         ## convert from coda to matrix  
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(exp(out[,x.cols]),2,quantile,c(0.025,0.5,0.975)) ## model was fit on log scale

plot(time,ci[2,],type='n',ylim=range(y,na.rm=TRUE),ylab="Flu Index",log='y',xlim=time[time.rng])
## adjust x-axis label to be monthly if zoomed
if(diff(time.rng) < 100){ 
  axis.Date(1, at=seq(time[time.rng[1]],time[time.rng[2]],by='month'), format = "%Y-%m")
}
ecoforecastR::ciEnvelope(time,ci[1,],ci[3,],col=ecoforecastR::col.alpha("lightBlue",0.75))
points(time,y,pch="+",cex=0.5)
```

### Next, lets look at the posterior distributions for `tau_add` and `tau_obs`, which we'll convert from precisions back into **standard deviations**.

```{r}
hist(1/sqrt(out[,1]),main=colnames(out)[1])
hist(1/sqrt(out[,2]),main=colnames(out)[2])
```

### We'll also want to look at the joint distribution of the two parameters to check whether the two parameters strongly covary.

```{r, fig.asp = 1.0}
plot(out[,1],out[,2],pch=".",xlab=colnames(out)[1],ylab=colnames(out)[2])
cor(out[,1:2])
```

## Part 3 - Back to one pond, but with temperature as a covariate

### Pull in the data we want to analyze

```{r}
YF_WaterTemp

# pulling out the dates here
time <- as.Date(YF_WaterTemp$`Month-Year`, "%m-%Y")

# Choose one pond to plot through time -- MP1
y = YF_WaterTemp %>%
  select(MP1)

temp = YF_AirTemp %>%
  select(Air_MonthAvg)
```

### Setting up the random walk

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t-1],tau_add)
    k[t-1] <- beta0 + m * temp[t-1]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

# Adding in temperature here for the Yakutat Forelands

### Setting up the random walk --> added in the air temperature here with only one pond

```{r}
RandomWalk_Temp = "
model{
  
  #### Data Model
  for(t in 1:n){
    y[t] ~ dnorm(x[t],tau_obs)
  }
  
  #### Process Model
  for(t in 2:n){
    x[t]~dnorm(k[t],tau_add)
    k[t] <- beta + delta * temp[t]
  }
  
  #### Priors
  x[1] ~ dnorm(x_ic,tau_ic)
  beta ~ dnorm(betap, sigmab)
  delta ~ dgamma(deltap, sigmad)
  tau_obs ~ dgamma(a_obs,r_obs)
  tau_add ~ dgamma(a_add,r_add)
}
"
```

### Next we need to define the data and priors as a list.

#for 'all' now

```{r}
data <- list(y=y$MP1,n=length(y),     ## data
             x_ic=0,tau_ic=0.2,      ## initial condition prior
             betap=0, sigmab=0.01,
             deltap=1, sigmad = 0.01,
             temp = temp, 
             a_obs=1,r_obs=1,           ## obs error prior
             a_add=1,r_add=1            ## process error prior
             )
```

### Next we need to definite the initial state of the model's parameters for each chain in the MCMC. The overall initialization is stored as a list the same length as the number of chains, where each chain is passed a list of the initial values for each parameter. Unlike the definition of the priors, which had to be done independent of the data, the initialization of the MCMC is allowed (and even encouraged) to use the data. However, each chain should be started from different initial conditions. We handle this below by basing the initial conditions for each chain off of a different random sample of the original data.

AW - take for each pond and have what was originally here with a sample from each pond and take a sample. Then average across ponds for each of these values. Also will indicate the importance between ponds by comparing this versus the average among ponds

```{r}
nchain = 3
init <- list()
for(i in 1:nchain){
  vect <- c(y$MP1)
  y.samp = sample(vect,length(vect),replace=TRUE)
  init[[i]] <- list(tau_add=1/var(diff(y.samp)),  ## initial guess on process precision
                    tau_obs=5/var(y.samp))        ## initial guess on obs precision
}
```

### Now that we've defined the model, the data, and the initialization, we need to send all this info to JAGS, which will return the JAGS model object.

```{r}
j.model   <- jags.model (file = textConnection(RandomWalk_Temp),
                             data = data,
                             inits = init,
                             n.chains = 3)
```

### Next, given the defined JAGS model, we'll want to take a few samples from the MCMC chain and assess when the model has converged. To take samples from the MCMC object we'll need to tell JAGS what variables to track and how many samples to take.

```{r, fig.asp = 1.0}
## burn-in
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("tau_add","tau_obs", "beta", "delta"),
                                n.iter = 1000)
plot(jags.out)
```

### Since rjags returns the samples as a CODA object, we can use any of the diagnostics in the R *coda* library to test for convergence, summarize the output, or visualize the chains.

Now that the model has converged we'll want to take a much larger sample from the MCMC and include the full vector of X's in the output

```{r}
jags.out   <- coda.samples (model = j.model,
                            variable.names = c("x","tau_add","tau_obs"),
                                n.iter = 10000)
```

### Given the full joint posterior samples, we're next going to visualize the output by just looking at the **95% credible interval of the time-series of X's** and compare that to the observed Y's. To do so we'll convert the coda output into a matrix and then calculate the quantiles. Looking at colnames(out) will show you that the first two columns are `tau_add` and `tau_obs`, so we calculate the CI starting from the 3rd column. We also transform the samples back from the log domain to the linear domain.

```{r}
time.rng = c(1,length(time))       ## adjust to zoom in and out
out <- as.matrix(jags.out)         ## convert from coda to matrix  
x.cols <- grep("^x",colnames(out)) ## grab all columns that start with the letter x
ci <- apply(exp(out[,x.cols]),2,quantile,c(0.025,0.5,0.975)) ## model was fit on log scale

plot(time,ci[2,],type='n',ylim=range(y,na.rm=TRUE),ylab="Flu Index",log='y',xlim=time[time.rng])
## adjust x-axis label to be monthly if zoomed
if(diff(time.rng) < 100){ 
  axis.Date(1, at=seq(time[time.rng[1]],time[time.rng[2]],by='month'), format = "%Y-%m")
}
ecoforecastR::ciEnvelope(time,ci[1,],ci[3,],col=ecoforecastR::col.alpha("lightBlue",0.75))
points(time,y,pch="+",cex=0.5)
```

###Next, lets look at the posterior distributions for `tau_add` and `tau_obs`, which we'll convert from precisions back into **standard deviations**.

```{r}
hist(1/sqrt(out[,1]),main=colnames(out)[1])
hist(1/sqrt(out[,2]),main=colnames(out)[2])
```

### We'll also want to look at the joint distribution of the two parameters to check whether the two parameters strongly covary.

```{r, fig.asp = 1.0}
plot(out[,1],out[,2],pch=".",xlab=colnames(out)[1],ylab=colnames(out)[2])
cor(out[,1:2])
```
