---
title: "CORDEX Data Summaries"
author: "Amaryllis Adey"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/amaryllisadey/Dropbox/SWEL/CopperRiverDelta/PondTemps/DataAnalysis/PondTempsAnalysis_GIT/")
```

```{r}
#Packages
library(dplyr)
library(ggpubr)
library(ggsci)
library(Rcpp)
library(reshape)
library(lubridate)
library(naniar)
library(gtools)
library(tidyr)
library(ncdf4)
library(raster)
library(sf)
library(tidync)
library(dplyr)
library(datetime)
```

Working with RCP 4.5 data

```{r}
#Read in all of the RCP 4.5 data
RCP45 <- tidync("/Users/amaryllisadey/Dropbox/SWEL/CopperRiverDelta/PondTemps/DataAnalysis/PondTempsAnalysis_GITNew/DataFiles/CORDEX/RCP_4.5/t_CORDEX-NAM_rcp45_mon_200601-210012.nc")

#Filter the data to the CRD
test <- RCP45

# Tell R what the target values are
CRD_target_lat <- 60.5437
CRD_target_lon <- 145.7260

# Filter this to the Lat Long
test_filtered <- test %>%
  hyper_filter(lat = between(abs(lat), CRD_target_lat-0.25, CRD_target_lat+0.25), 
               lon = between(abs(lon), CRD_target_lon-0.25, CRD_target_lon+0.25)) %>%
  hyper_tibble()

print(test_filtered)

CRD_45 <- test_filtered
rm(test_filtered)

# Attemp to change the cordex time
test <- CRD_45 %>%
  group_by(time) %>%
  summarize(avgTemp = mean(t), stdTemp = sd(t)) %>%
  ungroup() %>%
  mutate(date_time = as.Date(time)) 
test2 <- test %>%
  mutate(date_time = date_time %m-%years(120)) %>%
  ungroup() %>%
  dplyr::select(date_time, avgTemp, stdTemp)

CRD_45 <- test2

rm(test, test2)

CRD_45 <- data.frame(CRD_45)

#Filter the data to the YF
test <- RCP45

# Tell R what the target values are
YF_target_lat <- 59.5089
YF_target_lon <- 139.6599

# Filter this to Lat Long
test_filtered <- test %>%
  hyper_filter(lat = between(abs(lat), YF_target_lat-0.25, YF_target_lat+0.25), 
               lon = between(abs(lon), YF_target_lon-0.25, YF_target_lon+0.25)) %>%
  hyper_tibble()

print(test_filtered)

YF_45 <- test_filtered
rm(test_filtered)

# Attemp to change the cordex time
test <- YF_45 %>%
  group_by(time) %>%
  summarize(avgTemp = mean(t), stdTemp = sd(t)) %>%
  ungroup() %>%
  mutate(date_time = as.Date(time)) 
test2 <- test %>%
  mutate(date_time = date_time %m-%years(120)) %>%
  ungroup() %>%
  dplyr::select(date_time, avgTemp, stdTemp)

YF_45 <- test2

rm(test, test2)

YF_45 <- data.frame(YF_45)

```

Working with RCP 8.5 data

```{r}
#Read in all of the RCP 8.5 data
RCP85 <- tidync("/Users/amaryllisadey/Dropbox/SWEL/CopperRiverDelta/PondTemps/DataAnalysis/PondTempsAnalysis_GITNew/DataFiles/CORDEX/RCP_8.5/t_CORDEX-NAM_rcp45_mon_200601-210012.nc")

#Filter the data to the CRD
test <- RCP85

# Tell R what the target values are
CRD_target_lat <- 60.5437
CRD_target_lon <- 145.7260

# Filter this to the Lat Long
test_filtered <- test %>%
  hyper_filter(lat = between(abs(lat), CRD_target_lat-0.25, CRD_target_lat+0.25), 
               lon = between(abs(lon), CRD_target_lon-0.25, CRD_target_lon+0.25)) %>%
  hyper_tibble()

print(test_filtered)

CRD_85 <- test_filtered
rm(test_filtered)

data.frame(CRD_85)

# Attemp to change the cordex time
test <- CRD_85 %>%
  group_by(time) %>%
  summarize(avgTemp = mean(t), stdTemp = sd(t)) %>%
  ungroup() %>%
  mutate(date_time = as.Date(time)) 
test2 <- test %>%
  mutate(date_time = date_time %m-%years(120)) %>%
  ungroup() %>%
  dplyr::select(date_time, avgTemp, stdTemp)

CRD_85 <- test2

rm(test, test2)

CRD_85 <- data.frame(CRD_85)

#Filter the data to the YF
test <- RCP85

# Tell R what the target values are
YF_target_lat <- 59.5089
YF_target_lon <- 139.6599

# Filter this to Lat Long
test_filtered <- test %>%
  hyper_filter(lat = between(abs(lat), YF_target_lat-0.25, YF_target_lat+0.25), 
               lon = between(abs(lon), YF_target_lon-0.25, YF_target_lon+0.25)) %>%
  hyper_tibble()

print(test_filtered)

YF_85 <- test_filtered
rm(test_filtered)

# Attemp to change the cordex time
test <- YF_85 %>%
  group_by(time) %>%
  summarize(avgTemp = mean(t), stdTemp = sd(t)) %>%
  ungroup() %>%
  mutate(date_time = as.Date(time)) 
test2 <- test %>%
  mutate(date_time = date_time %m-% years(120)) %>%
  dplyr::select(date_time, avgTemp, stdTemp)

YF_85 <- test2

rm(test, test2)

YF_85 <- data.frame(YF_85)

```

Saving the files as .csv files that can be loaded into other code

```{r}
write.csv(CRD_45, "CORDEX_CRD_45.csv")
write.csv(CRD_85, "CORDEX_CRD_85.csv")

write.csv(YF_45, "CORDEX_YF_45.csv")
write.csv(YF_85, "CORDEX_YF_85.csv")

```

### Plot the forecasted data

#### Copper River Delta

```{r}
CRD_45

# Full time frame (2006-2100)
CRD_CORDEX <- ggplot(data = CRD_45, mapping = aes(x = date_time)) +
  geom_line(aes(y = avgTemp), linewidth = 0.2) +
  geom_ribbon(aes(ymax = (avgTemp + stdTemp), ymin = (avgTemp - stdTemp)))+
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Copper River CORDEX Forecast - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
CRD_CORDEX

ggsave("CRD_CORDEX_Forecast45.jpeg", plot = CRD_CORDEX, width = 8, height = 6)

# Limited time frame (2020-2100)
test <- CRD_45 %>%
  filter(date_time > "2019-12-01")

CRD_45 <- test

CRD_CORDEX <- ggplot(data = CRD_45, mapping = aes(x = date_time)) +
  geom_line(aes(y = avgTemp), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Copper River CORDEX Forecast - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
CRD_CORDEX

ggsave("CRD_CORDEX_Forecast4.5_2020.jpeg", plot = CRD_CORDEX, width = 8, height = 6)
```

#### Yakutat Forelands

```{r}
YF_45

# Full time frame (2006-2100)
YF_CORDEX <- ggplot(data = YF_45, mapping = aes(x = date_time)) +
  geom_line(aes(y = avgTemp), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Yakutat Forelands CORDEX Forecast - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
YF_CORDEX

ggsave("/YF_CORDEX_Forecast45.jpeg", plot = YF_CORDEX, width = 8, height = 6)

# Limited time frame (2020-2100)
test <- YF_45 %>%
  filter(date_time > "2019-12-01")

YF_45 <- test

YF_CORDEX <- ggplot(data = YF_45, mapping = aes(x = date_time)) +
  geom_line(aes(y = avgTemp), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Yakutat Forelands CORDEX Forecast - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
YF_CORDEX

ggsave("YF_CORDEX_Forecast4.5_2020.jpeg", plot = YF_CORDEX, width = 8, height = 6)
```

### Plot the forecasted data

#### Copper River Delta

```{r}
CRD_85

# Full time frame (2006-2100)
CRD_CORDEX <- ggplot(data = CRD_85, mapping = aes(x = date_time)) +
  geom_line(aes(y = avgTemp), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Copper River CORDEX Forecast - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
CRD_CORDEX

ggsave("CRD_CORDEX_Forecast85.jpeg", plot = CRD_CORDEX, width = 8, height = 6)

# Limited time frame (2020-2100)
test <- CRD_85 %>%
  filter(date_time > "2019-12-01")

CRD_85 <- test

CRD_CORDEX <- ggplot(data = CRD_85, mapping = aes(x = date_time)) +
  geom_line(aes(y = avgTemp), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Copper River CORDEX Forecast - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
CRD_CORDEX

ggsave("CRD_CORDEX_Forecast85_2020.jpeg", plot = CRD_CORDEX, width = 8, height = 6)
```

#### Yakutat Forelands

```{r}
YF_85

# Full time frame (2006-2100)
YF_CORDEX <- ggplot(data = YF_85, mapping = aes(x = date_time)) +
  geom_line(aes(y = avgTemp), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Yakutat Forelands CORDEX Forecast - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
YF_CORDEX

ggsave("YF_CORDEX_Forecast85.jpeg", plot = YF_CORDEX, width = 8, height = 6)

# Limited time frame (2020-2100)
test <- YF_85 %>%
  filter(date_time > "2019-12-01")

YF_85 <- test

YF_CORDEX <- ggplot(data = YF_85, mapping = aes(x = date_time)) +
  geom_line(aes(y = avgTemp), linewidth = 0.2) +
  xlab("Year") +
  ylab("Mean Forecast Temperature") +
  ggtitle("Yakutat Forelands CORDEX Forecast - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 
YF_CORDEX

ggsave("YF_CORDEX_Forecast85_2020.jpeg", plot = YF_CORDEX, width = 8, height = 6)
```

# Appending the Temperature Forecasts to the observed data

## Starting with RCP 4.5

### Creating a dataframe with missing values for all other aspects

Determining the columns needed for this

```{r}
CRD_ObsData <- read.csv("CR_PondAirTemp_mean.csv", header=TRUE)
head(CRD_ObsData)

CRD_ObsData <- data.frame(CRD_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CRD_ObsData <- CRD_ObsData %>%
  dplyr::rename(date=Date)
CRD_ObsData <- CRD_ObsData[,-1]
head(CRD_ObsData)

CRD_ObsData <- CRD_ObsData %>%
  dplyr::select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CRD_ObsData
```

Columns needed are date, Airport, Air_MonthAvg, Air_MonthMax, Air_MonthMin, Region, TIN, SQR, TIS, CAB, EYS, RHM, BVS, CAN, WDD

#### Copper River Delta

Checking the dimensions of the dataframe

```{r}
CRD_45
dim(CRD_45) # 972 rows here, so need to make 972 blank rows
str(CRD_45)

CRD_45$precision <- 1/((CRD_45$stdTemp)^2)

```

Creating an empty dataframe

```{r}
CRD_df <- CRD_ObsData[0,]
str(CRD_df)

attr(CRD_df, 'row.names') <- 1:nrow(CRD_45)

str(CRD_df)
```

#### Yakutat Forelands

Checking the dimensions of the dataframe

```{r}
YF_45
dim(YF_45) # 972 rows here, so need to make 972 blank rows
str(YF_45)
YF_45$precision <- 1/((YF_45$stdTemp)^2)
```

Checking the observational data

```{r}
YF_ObsData <- read.csv("DataFiles/YF_PondAirTemp_mean.csv", header=TRUE)
head(YF_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
YF_ObsData <- YF_ObsData %>%
  dplyr::rename(date=Date)
YF_ObsData <- YF_ObsData[,-1]
head(YF_ObsData)

YF_ObsData <- YF_ObsData %>%
  dplyr::select(date, Airport, Air_MonthAvg, std, precision, Region,  MP1, MP3, PL3, PL2, UBP3, MP5, MP8, UBP1, UBP2, PL1, UBP4)
YF_ObsData
```

Creating an empty dataframe

```{r}
YF_df <- YF_ObsData[0,]
str(YF_df)

attr(YF_df, 'row.names') <- 1:nrow(YF_45)

str(YF_df)
str(YF_ObsData)
```

### Adding in the data where it is available

#### Copper River Delta

```{r}
CRD_df

#values where there is data
CRD_df$date <- CRD_45$date
CRD_df$Air_MonthAvg <- CRD_45$avgTemp
CRD_df$std <- CRD_45$std
CRD_df$precision <- CRD_45$precision
CRD_df$Region <- rep("CRD", nrow(CRD_df))

#values where there is not data
CRD_df$Airport <- rep(NA, nrow(CRD_df))
CRD_df$BVS <- rep(NA, nrow(CRD_df))
CRD_df$CAB <- rep(NA, nrow(CRD_df))
CRD_df$CAN <- rep(NA, nrow(CRD_df))
CRD_df$EYS <- rep(NA, nrow(CRD_df))
CRD_df$RHM <- rep(NA, nrow(CRD_df))
CRD_df$SQR <- rep(NA, nrow(CRD_df))
CRD_df$TIN <- rep(NA, nrow(CRD_df))
CRD_df$TIS <- rep(NA, nrow(CRD_df))
CRD_df$WDD <- rep(NA, nrow(CRD_df))

str(CRD_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
CRD_df$date
CRD_ObsData$date <- as.Date(CRD_ObsData$date)

CRD_new_df <- rbind(CRD_ObsData, CRD_df)

str(CRD_new_df)

CRD_45Forecast <- CRD_new_df
CRD_45Forecast

#save this as a .csv file
write.csv(CRD_45Forecast, file = "CRD_45Forecast.csv")

```

#### Yakutat Forelands

Adding in the data where it is available

```{r}

#values where there is data
YF_df$date <- YF_45$date
YF_df$Air_MonthAvg <- YF_45$avgTemp
YF_df$std <- YF_45$std
YF_df$precision <- YF_45$precision
YF_df$Region <- rep("YF", nrow(YF_df))

#values where there is not data
YF_df$Airport <- rep(NA, nrow(YF_df))
YF_df$MP1<- rep(NA, nrow(YF_df))
YF_df$MP3 <- rep(NA, nrow(YF_df))
YF_df$MP5 <- rep(NA, nrow(YF_df))
YF_df$MP8 <- rep(NA, nrow(YF_df))
YF_df$PL1 <- rep(NA, nrow(YF_df))
YF_df$PL2 <- rep(NA, nrow(YF_df))
YF_df$PL3 <- rep(NA, nrow(YF_df))
YF_df$UBP1 <- rep(NA, nrow(YF_df))
YF_df$UBP2 <- rep(NA, nrow(YF_df))
YF_df$UBP3 <- rep(NA, nrow(YF_df))
YF_df$UBP4 <- rep(NA, nrow(YF_df))

str(YF_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
YF_df
YF_ObsData$date <- as.Date(YF_ObsData$date)

YF_new_df <- rbind(YF_ObsData, YF_df)

str(YF_new_df)

YF_45Forecast <- YF_new_df
YF_45Forecast

#save this as a .csv file
write.csv(YF_45Forecast, file = "YF_45Forecast.csv")

```

## Continuing with RCP 8.5

### Creating a dataframe with missing values for all other aspects

Determining the columns needed for this

```{r}
CRD_ObsData
```

Columns needed are date, Airport, Air_MonthAvg, Air_MonthMax, Air_MonthMin, Region, TIN, SQR, TIS, CAB, EYS, RHM, BVS, CAN, WDD

#### Copper River Delta

Checking the dimensions of the dataframe

```{r}
CRD_85
dim(CRD_85) # 972 rows here, so need to make 972 blank rows

CRD_85$precision <- 1/((CRD_85$stdTemp)^2) 
```

Checking the observational data

```{r}
CRD_ObsData
```

Creating an empty dataframe

```{r}
CRD_df <- CRD_ObsData[0,]
str(CRD_df)

attr(CRD_df, 'row.names') <- 1:nrow(CRD_85)

str(CRD_df)
```

#### Yakutat Forelands

Checking the dimensions of the dataframe

```{r}
YF_85
dim(YF_85) # 972 rows here, so need to make 972 blank rows

YF_85$precision <- 1/((YF_85$stdTemp)^2)
```

Checking the observational data

```{r}
YF_ObsData
```

Creating an empty dataframe

```{r}
YF_df <- YF_ObsData[0,]
str(YF_df)

attr(YF_df, 'row.names') <- 1:nrow(YF_85)

str(YF_df)
```

### Adding in the data where it is available

#### Copper River Delta

```{r}
CRD_df

#values where there is data
CRD_df$date <- CRD_85$date
CRD_df$Air_MonthAvg <- CRD_85$avgTemp
CRD_df$std <- CRD_85$std
CRD_df$precision <- CRD_85$precision
CRD_df$Region <- rep("CRD", nrow(CRD_df))

#values where there is not data
CRD_df$Airport <- rep(NA, nrow(CRD_df))
CRD_df$BVS <- rep(NA, nrow(CRD_df))
CRD_df$CAB <- rep(NA, nrow(CRD_df))
CRD_df$CAN <- rep(NA, nrow(CRD_df))
CRD_df$EYS <- rep(NA, nrow(CRD_df))
CRD_df$RHM <- rep(NA, nrow(CRD_df))
CRD_df$SQR <- rep(NA, nrow(CRD_df))
CRD_df$TIN <- rep(NA, nrow(CRD_df))
CRD_df$TIS <- rep(NA, nrow(CRD_df))
CRD_df$WDD <- rep(NA, nrow(CRD_df))

str(CRD_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
CRD_df
CRD_ObsData$date <- as.Date(CRD_ObsData$date)

CRD_new_df <- rbind(CRD_ObsData, CRD_df)

str(CRD_new_df)

CRD_85Forecast <- CRD_new_df
CRD_85

#save this as a .csv file
write.csv(CRD_85Forecast, file = "CRD_85Forecast.csv")

```

#### Yakutat Forelands

Adding in the data where it is available

```{r}

#values where there is data
YF_df$date <- YF_85$date
YF_df$Air_MonthAvg <- YF_85$avgTemp
YF_df$std <- YF_85$std
YF_df$precision <- YF_85$precision
YF_df$Region <- rep("YF", nrow(YF_df))

#values where there is not data
YF_df$Airport <- rep(NA, nrow(YF_df))
YF_df$MP1<- rep(NA, nrow(YF_df))
YF_df$MP3 <- rep(NA, nrow(YF_df))
YF_df$MP5 <- rep(NA, nrow(YF_df))
YF_df$MP8 <- rep(NA, nrow(YF_df))
YF_df$PL1 <- rep(NA, nrow(YF_df))
YF_df$PL2 <- rep(NA, nrow(YF_df))
YF_df$PL3 <- rep(NA, nrow(YF_df))
YF_df$UBP1 <- rep(NA, nrow(YF_df))
YF_df$UBP2 <- rep(NA, nrow(YF_df))
YF_df$UBP3 <- rep(NA, nrow(YF_df))
YF_df$UBP4 <- rep(NA, nrow(YF_df))

str(YF_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
YF_df
YF_ObsData$date <- as.Date(YF_ObsData$date)

YF_new_df <- rbind(YF_ObsData, YF_df)

str(YF_new_df)

YF_85Forecast <- YF_new_df
YF_85Forecast

#save this as a .csv file
write.csv(YF_85Forecast, file = "YF_85Forecast.csv")

```

# Attemping a quick plot of this data for both CR and YF

## RCP 4.5

### Copper River Delta

```{r}
CRD_45Forecast 

CRD_45Temps<- ggplot(data = CRD_45Forecast, mapping = aes(x = date)) +
  geom_ribbon(mapping = aes(ymin = (Air_MonthAvg - std), ymax = (Air_MonthAvg + std)), color = "grey") +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Copper River Delta Temperature - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
CRD_45Temps

ggsave("CRD_45Temps.jpeg", plot = CRD_45Temps, width = 8, height = 6)
```

### Yakutat Forelands

```{r}
YF_45Forecast 

YF_45Temps<- ggplot(data = YF_45Forecast, mapping = aes(x = date)) +
  geom_ribbon(mapping = aes(ymin = (Air_MonthAvg - std), ymax = (Air_MonthAvg + std)), color = "grey") +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Yakutat Forelands Temperature - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
YF_45Temps

ggsave("YF_45Temps.jpeg", plot = YF_45Temps, width = 8, height = 6)
```

## RCP 8.5

### Copper River Delta

```{r}
CRD_85Forecast 

CRD_85Temps<- ggplot(data = CRD_85Forecast, mapping = aes(x = date)) +
  geom_ribbon(mapping = aes(ymin = (Air_MonthAvg - std), ymax = (Air_MonthAvg + std)), color = "grey") +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Copper River Delta Temperature - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
CRD_85Temps

ggsave("CRD_85Temps.jpeg", plot = CRD_85Temps, width = 8, height = 6)
```

### Yakutat Forelands

```{r}
YF_85Forecast 

YF_85Temps<- ggplot(data = YF_85Forecast, mapping = aes(x = date)) +
  geom_ribbon(mapping = aes(ymin = (Air_MonthAvg - std), ymax = (Air_MonthAvg + std)), color = "grey") +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Yakutat Forelands Temperature - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
YF_85Temps

ggsave("YF_85Temps.jpeg", plot = YF_85Temps, width = 8, height = 6)
```
Bring in the SNAP data here to see how these CORDEX forecasts compare to the SNAP one since they seem much lower 

```{r}
CRD_snap45 <- read.csv() # file read in not working properly

CRD_45Forecast 

CRD_45Temps<- ggplot(data = CRD_45Forecast, mapping = aes(x = date)) +
  geom_ribbon(mapping = aes(ymin = (Air_MonthAvg - std), ymax = (Air_MonthAvg + std)), color = "grey") +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
  geom_line(data = CRD_snap45, mapping = aes(x = date, y = Air_MonthAvg, color = "red")) +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Copper River Delta Temperature - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
CRD_45Temps

ggsave("CRD_45Temps.jpeg", plot = CRD_45Temps, width = 8, height = 6)

```



