---
title: "Pond Temps - Delta Graphs and Seasonal Trends"
output: html_notebook
---
This is code used to make plots based on output from the "IndividualPondModels_Forecasts85" for boththe CR and YF
Read in the output files for each pond/RCP combination then work on delta figures and figures like Hamlet et al. 2020

```{r}
rm(list = ls())

library(tidyverse)
library(dplyr)
library(rjags)
library(ggplot2)

set.seed(1)
```

Starting with the Yakutat Forelands

RCP 85

# Reading in the data
```{r}
# Load required libraries
library(ggplot2)
library(dplyr)
library(lubridate)
library(ggpubr)

# Reading in the data
Mod_YF85 <- read.csv("Corr_Mod_YF85.csv", header = TRUE)

# Convert 'time' column to Date format
Mod_YF85$time <- as.Date(Mod_YF85$time, format = "%Y-%m-%d")

# Extract year and month information
Mod_YF85 <- Mod_YF85 %>%
  mutate(Year = year(time),
         Month = format(time, "%B"))

```

# Making plots of error by season

## 2012-2020
```{r}
# Extract unique pond names
short_names <- unique(Mod_YF85$pond_name)

# Initialize an empty dataframe to store monthly mean results
monthly_means <- data.frame(Pond = character(), Month = character(), AvgTemp = numeric(), Year = numeric(), stringsAsFactors = FALSE)

# Loop through each pond
for (short_name in short_names) {
  # Filter data for the current pond
  pond_data <- subset(Mod_YF85, pond_name == short_name & Year >= 2012 & Year <= 2020)
  
  # Loop through each month
  for (month in month.name) {
    # Filter data for the current month
    filtered_data <- subset(pond_data, Month == month)
    
    # Calculate the mean of 'med' for the current month
    avg_med <- mean(filtered_data$med, na.rm = TRUE)
    
    # Append results to monthly_means dataframe
    monthly_means <- rbind(monthly_means, data.frame(Pond = short_name, Month = month, AvgTemp = avg_med, Year = 2020))
  }
}

# Plotting the data
ggplot(monthly_means, aes(x = Month, y = AvgTemp, color = Pond, group = Pond)) +
  geom_line() +
  geom_point() +
  scale_x_discrete(limits = month.name) +
  labs(title = "Monthly Average Temperatures for Each Pond (2012–2020)",
       x = "Month",
       y = "Average Temperature (°C)",
       color = "Pond") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## 2030, 2060, and 2090
```{r}
# Define the correct order for months
month_order <- c("January", "February", "March", "April", "May", "June", 
                 "July", "August", "September", "October", "November", "December")

# Function to calculate monthly averages by decade
calculate_monthly_averages <- function(data, decade_start, decade_end) {
  monthly_avg_table <- data.frame(Pond = character(), Month = character(), AvgTemp = numeric(), Year = integer(), stringsAsFactors = FALSE)
  
  for (yr in decade_start:decade_end) {
    yearly_data <- data %>% filter(Year == yr)
    
    for (pond_name in unique(data$pond_name)) {
      pond_data <- yearly_data %>% filter(pond_name == pond_name)
      
      for (month in month_order) {
        monthly_data <- pond_data %>% filter(format(time, "%B") == month)
        
        avg_temp <- mean(monthly_data$med, na.rm = TRUE)
        
        monthly_avg_table <- rbind(monthly_avg_table, data.frame(
          Pond = pond_name,
          Month = month,
          AvgTemp = avg_temp,
          Year = yr
        ))
      }
    }
  }
  return(monthly_avg_table)
}

# Calculate averages for each decade
monthly_avgX50_table_2030s <- calculate_monthly_averages(Mod_YF85, 2030, 2039)
monthly_avgX50_table_2060s <- calculate_monthly_averages(Mod_YF85, 2060, 2069)
monthly_avgX50_table_2090s <- calculate_monthly_averages(Mod_YF85, 2090, 2099)

```
# Combine all the monthly average tables into one dataframe 

```{r}
# Add a Decade column to each dataframe
monthly_means <- monthly_means %>% mutate(Decade = "2020s")
monthly_avgX50_table_2030s <- monthly_avgX50_table_2030s %>% mutate(Decade = "2030s")
monthly_avgX50_table_2060s <- monthly_avgX50_table_2060s %>% mutate(Decade = "2060s")
monthly_avgX50_table_2090s <- monthly_avgX50_table_2090s %>% mutate(Decade = "2090s")

# Combine all data into one dataframe
combined_monthly_avg <- bind_rows(
  monthly_means,
  monthly_avgX50_table_2030s,
  monthly_avgX50_table_2060s,
  monthly_avgX50_table_2090s
)

# Reference year data
ref_year_avg <- combined_monthly_avg %>%
  filter(Year == 2020) %>%
  group_by(Month) %>%
  summarise(avg_temp_ref_year = mean(AvgTemp, na.rm = TRUE))

# Calculate differences
combined_monthly_avg_with_diff <- combined_monthly_avg %>%
  left_join(ref_year_avg, by = "Month") %>%
  mutate(diff_AvgTemp = AvgTemp - avg_temp_ref_year) %>%
  filter(!is.na(diff_AvgTemp))  # Remove rows with missing differences


```

# Calculate the differences here as compared to 2020
```{r}
# Group by month and decade, calculate statistics
summary_stats2 <- combined_monthly_avg_with_diff %>%
  mutate(Month = factor(Month, levels = month_order)) %>%
  group_by(Month, Decade) %>%
  summarise(
    mean_diff_AvgTemp = mean(diff_AvgTemp, na.rm = TRUE),
    sd_diff_AvgTemp = sd(diff_AvgTemp, na.rm = TRUE)
  )

# Function to create decade plots
create_decade_plot <- function(decade) {
  data <- summary_stats2 %>% filter(Decade == decade)
  
  ggplot(data, aes(x = as.numeric(Month))) +
    geom_line(aes(y = mean_diff_AvgTemp)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
    geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp),
                fill = "blue", alpha = 0.3) +
    scale_x_continuous(breaks = 1:12, labels = month.abb) +
    ylim(-4, 6) +
    theme_classic() +
    labs(title = paste("Delta Temp:", decade),
         x = "", y = "Delta Temp (°C)") +
    theme(axis.text.x = element_text(angle = 65, hjust = 1))
}

# Generate plots
MonthlyStats_2030Plot <- create_decade_plot("2030s")
MonthlyStats_2060Plot <- create_decade_plot("2060s")
MonthlyStats_2090Plot <- create_decade_plot("2090s")

# Combine plots
DeltaMonth_YF85_plot <- ggarrange(MonthlyStats_2030Plot, MonthlyStats_2060Plot, MonthlyStats_2090Plot, ncol = 3)
DeltaMonth_YF85_plot

# Save plot
ggsave("Corr_DeltaMonthYF85.png", plot = DeltaMonth_YF85_plot, width = 12, height = 3)

```


Combine the plots together here for now
```{r}
library(ggpubr)

DeltaMonth_YF85_plot <- ggarrange(plot_2030s, plot_2060s, plot_2090s, ncol = 3)
ggsave("Corr_DeltaMonthYF85.png", plot = DeltaMonth_YF85_plot, width = 12, height = 4)

write.csv(summary_stats %>% filter(Decade == "2030s"), "Corr_MonthlyStatsYF85_2030.csv")
write.csv(summary_stats %>% filter(Decade == "2060s"), "Corr_MonthlyStatsYF85_2060.csv")
write.csv(summary_stats %>% filter(Decade == "2090s"), "Corr_MonthlyStatsYF85_2090.csv")

```

