---
title: "Pond Temps - Delta Graphs and Seasonal Trends"
output: html_notebook
---
This is code used to make plots based on output from the "IndividualPondModels_Forecasts85" for boththe CR and YF
Read in the output files for each pond/RCP combination then work on delta figures and figures like Hamlet et al. 2020

```{r}
rm(list = ls())

library(tidyverse)
library(dplyr)
library(rjags)
library(ggplot2)

set.seed(1)
```

Starting with the Copper River Delta

RCP 85

# Reading in the data
```{r}
Mod_CRD85 <- read.csv("Corr_Mod_CRD85.csv", header = T)
```

 
### 2012-2020
```{r}
# List of unique pond names from the dataframe
pond_names <- unique(Mod_CRD85$pond_name)

# Initialize an empty dataframe to store results
avgMed_2020 <- data.frame(Pond = pond_names, AvgMed_2020 = numeric(length(pond_names)))

# Loop through each pond name
for (i in seq_along(pond_names)) {
  pond_name <- pond_names[i]
  
  # Filter data for the current pond and for dates before January 1, 2021
  filtered_data <- Mod_CRD85 %>%
    filter(pond_name == !!pond_name, as.Date(time) < as.Date("2021-01-01"))
  
  # Calculate the mean of 'med' (which corresponds to X50%)
  avg_med <- mean(filtered_data$med, na.rm = TRUE)
  
  # Store the result in the dataframe
  avgMed_2020[i, "AvgMed_2020"] <- avg_med
}

# Print the dataframe
print(avgMed_2020)
```

### 2030-2039
```{r}
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)

# List of unique pond names from the dataframe
pond_names <- unique(Mod_CRD85$pond_name)

# Initialize a dataframe to store results with years as columns
avgMed_2030 <- data.frame(Pond = pond_names)

# Loop through each year from 2030 to 2039
for (yr in 2030:2039) {
  # Calculate average medians for each year and store as a new column
  avgMed_col <- sapply(pond_names, function(pond_name) {
    yearly_data <- Mod_CRD85 %>%
      filter(pond_name == !!pond_name, year(as.Date(time)) == yr)
    
    avg_med <- mean(yearly_data$med, na.rm = TRUE)
    return(avg_med)
  })
  
  # Add the results as a new column to the dataframe
  avgMed_2030[[as.character(yr)]] <- avgMed_col
}

# Print the final dataframe
print(avgMed_2030)

# Convert to long format for ggplot2
avgMed_2030_long <- avgMed_2030 %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "AvgMed")

# Plot the data
ggplot(avgMed_2030_long, aes(x = Year, y = AvgMed, color = Pond, group = Pond)) +
  geom_line() +
  geom_point() +
  labs(title = "Average Median per Pond from 2030 to 2039",
       x = "Year",
       y = "Average Median") +
  theme_minimal()
```

### 2060-2069
```{r}
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)

# List of unique pond names from the dataframe
pond_names <- unique(Mod_CRD85$pond_name)

# Initialize a dataframe to store results with years as columns
avgMed_2060 <- data.frame(Pond = pond_names)

# Loop through each year from 2060 to 2069
for (yr in 2060:2069) {
  # Calculate average medians for each year and store as a new column
  avgMed_col <- sapply(pond_names, function(pond_name) {
    yearly_data <- Mod_CRD85 %>%
      filter(pond_name == !!pond_name, year(as.Date(time)) == yr)
    
    avg_med <- mean(yearly_data$med, na.rm = TRUE)
    return(avg_med)
  })
  
  # Add the results as a new column to the dataframe
  avgMed_2060[[as.character(yr)]] <- avgMed_col
}

# Print the final dataframe
print(avgMed_2060)

# Convert to long format for ggplot2
avgMed_2060_long <- avgMed_2060 %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "AvgMed")

# Plot the data
ggplot(avgMed_2060_long, aes(x = Year, y = AvgMed, color = Pond, group = Pond)) +
  geom_line() +
  geom_point() +
  labs(title = "Average Median per Pond from 2060 to 2069",
       x = "Year",
       y = "Average Median") +
  theme_minimal()
```

### 2090-2099
```{r}
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)

# List of unique pond names from the dataframe
pond_names <- unique(Mod_CRD85$pond_name)

# Initialize a dataframe to store results with years as columns
avgMed_2090 <- data.frame(Pond = pond_names)

# Loop through each year from 2090 to 2099
for (yr in 2090:2099) {
  # Calculate average medians for each year and store as a new column
  avgMed_col <- sapply(pond_names, function(pond_name) {
    yearly_data <- Mod_CRD85 %>%
      filter(pond_name == !!pond_name, year(as.Date(time)) == yr)
    
    avg_med <- mean(yearly_data$med, na.rm = TRUE)
    return(avg_med)
  })
  
  # Add the results as a new column to the dataframe
  avgMed_2090[[as.character(yr)]] <- avgMed_col
}

# Print the final dataframe
print(avgMed_2090)

# Convert to long format for ggplot2
avgMed_2090_long <- avgMed_2090 %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "AvgMed")

# Plot the data
ggplot(avgMed_2090_long, aes(x = Year, y = AvgMed, color = Pond, group = Pond)) +
  geom_line() +
  geom_point() +
  labs(title = "Average Median per Pond from 2090 to 2099",
       x = "Year",
       y = "Average Median") +
  theme_minimal()
```
```{r}
# Rename columns to reflect only the years
avgMed_2020 <- avgMed_2020 %>%
  rename(AvgMed_2020 = AvgMed_2020)

avgMed_2030 <- avgMed_2030 %>%
  rename(`2030` = `2030`)

avgMed_2060 <- avgMed_2060 %>%
  rename(`2060` = `2060`)

avgMed_2090 <- avgMed_2090 %>%
  rename(`2090` = `2090`)

# Combine the data frames
combined_avgMed <- avgMed_2020 %>%
  left_join(avgMed_2030, by = "Pond") %>%
  left_join(avgMed_2060, by = "Pond") %>%
  left_join(avgMed_2090, by = "Pond")

# Print the final dataframe
print(combined_avgMed)

# Convert to long format for ggplot2, excluding AvgMed_2020
avgMed_long <- combined_avgMed %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "AvgMed") %>%
  filter(Year %in% as.character(2021:2199))

# Plot the data
ggplot(avgMed_long, aes(x = Year, y = AvgMed, color = Pond, group = Pond)) +
  geom_line() +
  geom_point() +
  labs(title = "Average Median per Pond in the CRD - RCP 8.5",
       x = "Year",
       y = "Average Median") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),# Rotate x-axis labels by 45 degrees
        plot.title = element_text(hjust = 0.5))  
```


# Calcuate the differences here as new columns in the dataframe

```{r}
library(dplyr)

# Assuming combined_avgMed is loaded and structure verified
Delta_combined_avgMed <- combined_avgMed  # Use your actual dataframe

# Columns to calculate differences for
years <- c("2030", "2031", "2032", "2033", "2034", 
           "2035", "2036", "2037", "2038", "2039", 
           "2060", "2061", "2062", "2063", "2064", 
           "2065", "2066", "2067", "2068", "2069", 
           "2090", "2091", "2092", "2093", "2094", 
           "2095", "2096", "2097", "2098", "2099")

# Loop through each year and calculate the difference
for (year in years) {
  # Create a new column for the difference
  diff_col_name <- paste0("diff_", year)
  Delta_combined_avgMed[[diff_col_name]] <- Delta_combined_avgMed[[year]] - Delta_combined_avgMed$`AvgMed_2020`
}

# Display the structure of Delta_combined_avgMed to verify
str(Delta_combined_avgMed)
```

```{r}
# Create a new dataframe with the calculated means and standard deviations
Delta_CRD85 <- Delta_combined_avgMed %>%
  mutate(
    Mean_2030_2039 = rowMeans(select(., starts_with("diff_203")), na.rm = TRUE),  # Calculate row-wise mean for columns starting with "diff_203"
    SD_2030_2039 = rowSds(as.matrix(select(., starts_with("diff_203"))), na.rm = TRUE),  # Calculate row-wise SD for columns starting with "diff_203"
    Mean_2060_2069 = rowMeans(select(., starts_with("diff_206")), na.rm = TRUE),  # Calculate row-wise mean for columns starting with "diff_206"
    SD_2060_2069 = rowSds(as.matrix(select(., starts_with("diff_206"))), na.rm = TRUE),  # Calculate row-wise SD for columns starting with "diff_206"
    Mean_2090_2099 = rowMeans(select(., starts_with("diff_209")), na.rm = TRUE),  # Calculate row-wise mean for columns starting with "diff_209"
    SD_2090_2099 = rowSds(as.matrix(select(., starts_with("diff_209"))), na.rm = TRUE)  # Calculate row-wise SD for columns starting with "diff_209"
  )

# Print the structure of the new dataframe
str(Delta_CRD85)

# Optionally, view the first few rows to check the new columns
head(Delta_CRD85)
```

## Put these into total mean/standard deviations

```{r}
# Load necessary libraries
library(dplyr)

# Select and rename columns from Delta_CRD85
CRD85_decadal <- Delta_CRD85 %>%
  select(Pond, starts_with("Mean_"), starts_with("SD_"))

# Rename columns without the 'Y' prefix for means
names(CRD85_decadal) <- sub("^Mean_", "Mean_", names(CRD85_decadal))

# Rename columns without the 'Y' prefix for standard deviations
names(CRD85_decadal) <- sub("^SD_", "SD_", names(CRD85_decadal))

# Print the structure of the updated dataframe
str(CRD85_decadal)

# Optionally, view the first few rows to check the new columns
head(CRD85_decadal)
```

# Making the plots of these deltas
```{r}
# Load required packages
library(ggplot2)
library(gridExtra)
library(ggpubr)

# Create a grouped bar plot with error bars for Mean 2030
plot_2030 <- ggplot(CRD85_decadal, aes(x = Pond, y = Mean_2030_2039)) +
  geom_bar(stat = "identity", fill = "#FFD92F", color = "black") +
  geom_errorbar(aes(ymin = Mean_2030_2039 - SD_2030_2039, ymax = Mean_2030_2039 + SD_2030_2039),
                width = 0.4,  # Adjust the width of error bars as needed
                position = position_dodge(width = 0.9)) +
  labs(x = "", y = "Delta Temp (C)") +
  ylim(-1, 5) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))  # Centered title
plot_2030

# Create a grouped bar plot with error bars for Mean 2060
plot_2060 <- ggplot(CRD85_decadal, aes(x = Pond, y = Mean_2060_2069)) +
  geom_bar(stat = "identity", fill = "#F46D43", color = "black") +
  geom_errorbar(aes(ymin = Mean_2060_2069 - SD_2060_2069, ymax = Mean_2060_2069 + SD_2060_2069),
                width = 0.4,  # Adjust the width of error bars as needed
                position = position_dodge(width = 0.9)) +
  labs(x = "", y = "") +  # No y-axis label
  ylim(-1, 5) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) +  # Centered title 
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())   # Remove Y-axis ticks
plot_2060

# Create a grouped bar plot with error bars for Mean 2090
plot_2090 <- ggplot(CRD85_decadal, aes(x = Pond, y = Mean_2090_2099)) +
  geom_bar(stat = "identity", fill = "#D73027", color = "black") +
  geom_errorbar(aes(ymin = Mean_2090_2099 - SD_2090_2099, ymax = Mean_2090_2099 + SD_2090_2099),
                width = 0.4,  # Adjust the width of error bars as needed
                position = position_dodge(width = 0.9)) +
  labs(x = "", y = "") +  # No y-axis label
  ylim(-1, 5) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) +  # Centered title 
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())   # Remove Y-axis ticks
plot_2090

# Combine plots into a panel plot
Delta_CRD85_plot <- ggarrange(plot_2030, plot_2060, plot_2090, ncol = 3)

# Save the combined plot
ggsave("Corr_DeltaCRD85.png", plot = Delta_CRD85_plot, width = 8, height = 3)
```

# Statistical tests for each of these ponds for 2030s, 2060s, and 2090s

### 2030s
```{r}
Delta_CRD85

# Load required libraries
library(dplyr)
library(broom)
library(tidyr)

# Step 1: Prepare the data
data_long <- Delta_CRD85 %>%
  select(Pond, matches("^diff_203[0-9]$")) %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "Difference")

boxplot(data_long$Difference ~ data_long$Pond)

# Step 2: Conduct ANOVA by grouping by Pond
res_CRD85 <- aov(Difference ~ Pond, data = data_long)
summary(res_CRD85)

# Step 3: Tukey HSD
post_test_CRD85 <- TukeyHSD(res_CRD85)
post_test_CRD85

```

### 2060s
```{r}
Delta_CRD85

# Load required libraries
library(dplyr)
library(broom)

# Step 1: Prepare the data
data_long <- Delta_CRD85 %>%
  select(Pond, matches("^diff_206[0-9]$")) %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "Difference")

boxplot(data_long$Difference ~ data_long$Pond)

# Step 2: Conduct ANOVA by grouping by Pond
res_CRD85 <- aov(Difference ~ Pond, data = data_long)
summary(res_CRD85)

# Step 3: Tukey HSD
post_test_CRD85 <- TukeyHSD(res_CRD85)
post_test_CRD85
```

### 2090s
```{r}
Delta_CRD85

# Load required libraries
library(dplyr)
library(broom)

# Step 1: Prepare the data
data_long <- Delta_CRD85 %>%
  select(Pond, matches("^diff_209[0-9]$")) %>%
  pivot_longer(cols = -Pond, names_to = "Year", values_to = "Difference")

boxplot(data_long$Difference ~ data_long$Pond)

# Step 2: Conduct ANOVA by grouping by Pond
res_CRD85 <- aov(Difference ~ Pond, data = data_long)
summary(res_CRD85)

# Step 3: Tukey HSD
post_test_CRD85 <- TukeyHSD(res_CRD85)
post_test_CRD85
```

# Making plots of error by season

## 2012-2020
```{r}
# Extract unique pond names from the dataframe
short_names <- unique(Mod_CRD85$pond_name)

# Initialize an empty dataframe to store monthly mean results
monthly_means <- data.frame(Pond = character(), Month = character(), AvgTemp = numeric(), Year = numeric(), stringsAsFactors = FALSE)

# Convert 'time' column to Date class
Mod_CRD85$time <- as.Date(Mod_CRD85$time, format="%Y-%m-%d")

# Extract year and month from 'time' column
Mod_CRD85$Year <- format(Mod_CRD85$time, "%Y")
Mod_CRD85$Month <- format(Mod_CRD85$time, "%B")

# Loop through each unique pond name
for (short_name in short_names) {
  # Filter data for the current pond
  pond_data <- subset(Mod_CRD85, pond_name == short_name)
  
  # Loop through each month
  for (month in month.name) {
    # Filter data for the current month
    filtered_data <- subset(pond_data, Month == month)
    
    # Calculate the mean of 'med' for the current month
    avg_med <- mean(filtered_data$med, na.rm = TRUE)
    
    # Append results to monthly_means dataframe
    monthly_means <- rbind(monthly_means, data.frame(Pond = short_name, Month = month, AvgTemp = avg_med, Year = 2020))
  }
}

# Print the monthly means dataframe
print(monthly_means)

# Plotting the data
ggplot(monthly_means, aes(x = Month, y = AvgTemp, color = Pond, group = Pond)) +
  geom_line() +
  geom_point() +
  scale_x_discrete(limits = month.name) +
  labs(title = "Monthly Average Temperatures for Each Pond",
       x = "Month",
       y = "Average Temperature",
       color = "Pond") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 2030, 2060, and 2090
```{r}
# Define the correct order for months
month_order <- c("January", "February", "March", "April", "May", "June", 
                  "July", "August", "September", "October", "November", "December")

# Recalculate monthly averages with correct month ordering
calculate_monthly_averages <- function(data, decade_start, decade_end) {
  # Initialize empty dataframe to store monthly averages
  monthly_avg_table <- data.frame(Pond = character(), Month = factor(), AvgTemp = numeric(), Year = integer(), stringsAsFactors = FALSE)
  
  # Loop through each year in the decade
  for (yr in decade_start:decade_end) {
    # Filter data for the current year
    yearly_data <- data %>%
      filter(year(time) == yr)
    
    # Loop through each pond
    for (pond_name in unique(data$pond_name)) {
      pond_data <- yearly_data %>%
        filter(pond_name == pond_name)
      
      # Loop through each month
      for (month in month_order) {
        # Filter data for the current month
        monthly_data <- pond_data %>%
          filter(format(time, "%B") == month)
        
        # Calculate the mean of 'med' for the current month
        avg_temp <- mean(monthly_data$med, na.rm = TRUE)
        
        # Append results to monthly_avg_table dataframe
        monthly_avg_table <- rbind(monthly_avg_table, data.frame(Pond = pond_name, Month = factor(month, levels = month_order), AvgTemp = avg_temp, Year = yr))
      }
    }
  }
  
  return(monthly_avg_table)
}

# Calculate monthly averages for each decade
monthly_avgX50_table_2030s <- calculate_monthly_averages(Mod_CRD85, 2030, 2039)
monthly_avgX50_table_2060s <- calculate_monthly_averages(Mod_CRD85, 2060, 2069)
monthly_avgX50_table_2090s <- calculate_monthly_averages(Mod_CRD85, 2090, 2099)
```
# Combine all the monthly average tables into one dataframe 

```{r}
# Add a Decade column to each dataframe
monthly_means <- monthly_means %>%
  mutate(Decade = "2020s")

monthly_avgX50_table_2030s <- monthly_avgX50_table_2030s %>%
  mutate(Decade = "2030s")

monthly_avgX50_table_2060s <- monthly_avgX50_table_2060s %>%
  mutate(Decade = "2060s")

monthly_avgX50_table_2090s <- monthly_avgX50_table_2090s %>%
  mutate(Decade = "2090s")

# Combine all monthly average tables into one dataframe
combined_monthly_avg <- bind_rows(
  monthly_means,
  monthly_avgX50_table_2030s,
  monthly_avgX50_table_2060s,
  monthly_avgX50_table_2090s
)

# Print the combined dataframe
print(combined_monthly_avg)
```

# Calculate the differences here as compared to 2020
```{r}
# Define the reference year
reference_year <- 2020

# Check if the column exists and is correctly populated
summary(combined_monthly_avg)

# Calculate average temperature for each month in the reference year
ref_year_avg <- combined_monthly_avg %>%
  filter(Year == reference_year) %>%
  group_by(Month) %>%
  summarise(avg_temp_ref_year = mean(AvgTemp, na.rm = TRUE))

# Join reference year averages with the main dataset
combined_monthly_avg_with_diff <- combined_monthly_avg %>%
  left_join(ref_year_avg, by = "Month") %>%
  mutate(diff_AvgTemp = AvgTemp - avg_temp_ref_year) %>%
  select(-avg_temp_ref_year)  # Remove the reference year average column

# Display the structure to verify
str(combined_monthly_avg_with_diff)

# Filter out data for the reference year
DeltaMonthly_CRD85 <- combined_monthly_avg_with_diff %>%
  filter(Year != reference_year)

# Display the filtered data
print(DeltaMonthly_CRD85)

```


# Calculate mean differences

```{r}
library(dplyr)

DeltaMonthly_CRD85

DiffMonthly_CRD85 <- DeltaMonthly_CRD85 %>%
  select(Pond, Month, Year, diff_AvgTemp)

# Create a new column based on year ranges
DiffMonthly_CRD85$Decade <- ifelse(DiffMonthly_CRD85$Year >= 2030 & DiffMonthly_CRD85$Year < 2040, "2030s",
                          ifelse(DiffMonthly_CRD85$Year >= 2060 & DiffMonthly_CRD85$Year < 2070, "2060s",
                          ifelse(DiffMonthly_CRD85$Year >= 2090 & DiffMonthly_CRD85$Year < 2100, "2090s",
                          "Other")))

# Print the updated data frame
print(DiffMonthly_CRD85)
  
# Convert Month to Date type
DiffMonthly_CRD85 <- DiffMonthly_CRD85 %>%
  mutate(Month = as.Date(paste0(Month, " 1 ", Year), format = "%B %d %Y"))

print(DiffMonthly_CRD85)

# Group by Pond and Month, then calculate mean and standard deviation of diff_AvgTemp
summary_stats <- DiffMonthly_CRD85 %>%
  dplyr::mutate(Month = format(Month, "%m")) %>%  # Extracts YYYY-MM format
  dplyr::group_by(Pond, Month, Decade) %>%
  dplyr::summarise(
    mean_diff_AvgTemp = mean(diff_AvgTemp, na.rm = TRUE),
    sd_diff_AvgTemp = sd(diff_AvgTemp, na.rm = TRUE)
  )

# View the summary statistics
print(summary_stats)

# Group by just Month, then calculate mean and standard deviation of diff_AvgTemp
summary_stats2 <- DiffMonthly_CRD85 %>%
  dplyr::mutate(Month = format(Month, "%m")) %>%  # Extracts YYYY-MM format
  dplyr::group_by(Month, Decade) %>%
  dplyr::summarise(
    mean_diff_AvgTemp = mean(diff_AvgTemp, na.rm = TRUE),
    sd_diff_AvgTemp = sd(diff_AvgTemp, na.rm = TRUE)
  )

# View the summary statistics
print(summary_stats2)
```

Plotting the monthly temperature change graphs

2030s
```{r}
summary_stats # grouped by month and pond
summary_stats2 # grouped by just month

# 2. Monthly averages across ponds
MonthlyStats_2030 <- summary_stats2 %>%
  filter(Decade == "2030s") 

ggplot(data = MonthlyStats_2030, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStats_2030$Month <- as.numeric(MonthlyStats_2030$Month)

# Plot using ggplot
MonthlyStats_2030Plot <- ggplot(data = MonthlyStats_2030, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-4,6)+
  theme_classic() +
  labs(x = "", y = "Delta Temp (C)") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
MonthlyStats_2030Plot
```

2060s
```{r}
# 2. Monthly averages across ponds
MonthlyStats_2060 <- summary_stats2 %>%
  filter(Decade == "2060s") 

ggplot(data = MonthlyStats_2060, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStats_2060$Month <- as.numeric(MonthlyStats_2060$Month)

# Plot using ggplot
MonthlyStats_2060Plot <- ggplot(data = MonthlyStats_2060, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-4,6)+
  theme_classic() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStats_2060Plot
```
2090s
```{r}
# 2. Monthly averages across ponds
MonthlyStats_2090 <- summary_stats2 %>%
  filter(Decade == "2090s") 

ggplot(data = MonthlyStats_2090, aes(x = Month)) +
  geom_point(aes(y = mean_diff_AvgTemp)) +
  theme_classic()

# Convert Month from character to numeric to ensure proper ordering in ggplot
MonthlyStats_2090$Month <- as.numeric(MonthlyStats_2090$Month)

# Plot using ggplot
MonthlyStats_2090Plot <- ggplot(data = MonthlyStats_2090, aes(x = Month)) +
  geom_line(aes(y = mean_diff_AvgTemp)) +  # Point plot with mean values
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal dashed line at y = 0
  geom_ribbon(aes(ymin = mean_diff_AvgTemp - sd_diff_AvgTemp, ymax = mean_diff_AvgTemp + sd_diff_AvgTemp), fill = "blue", alpha = 0.3) +  # Ribbon for positive and negative error based on sd values
  scale_x_continuous(breaks = 1:12, labels = month.abb) +  # Adjust x-axis labels to show month abbreviations
  ylim(-4,6)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1)) +
  labs(x = "", y = "") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))+
  theme(axis.text.y = element_blank(),  # Remove Y-axis tick mark labels
        axis.ticks.y = element_blank())
MonthlyStats_2090Plot
```

Combine the plots together here for now
```{r}
# Combine plots into a panel plot
DeltaMonth_CRD85_plot <- ggarrange(MonthlyStats_2030Plot, MonthlyStats_2060Plot, MonthlyStats_2090Plot, ncol = 3)

# Save the combined plot
ggsave("Corr_DeltaMonthCRD85.png", plot = DeltaMonth_CRD85_plot, width = 12, height = 3)

```

Create .csv files of the data for these graphs
```{r}
write.csv(MonthlyStats_2030, file = "Corr_MonthlyStatsCRD85_2030.csv")
write.csv(MonthlyStats_2060, file = "Corr_MonthlyStatsCRD85_2060.csv")
write.csv(MonthlyStats_2090, file = "Corr_MonthlyStatsCRD85_2090.csv")
```
