---
title: "SNAP_RCP_DataExtractionOrganization"
author: "Amaryllis Adey"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setting up the workspace

```{r}
library(dplyr)
library(ggplot2)

setwd("~/Library/CloudStorage/Dropbox/SWEL/CopperRiverDelta/PondTemps/DataAnalysis/PondTempsAnalysis")
```

# Changing this to be reading in the bias corrected forecast data instead

## Read in the data

```{r}
SNAP_CR_45 <- read.csv("BiasCorrection/Geo_BiasCorrection/DataFiles/SNAP_CR45_corrected_monthly.csv")
SNAP_CR_45$date <- as.Date(SNAP_CR_45$date,format="%Y-%m-%d")
SNAP_CR_45
SNAP_CR_85 <- read.csv("BiasCorrection/Geo_BiasCorrection/DataFiles/SNAP_CR85_corrected_monthly.csv")
SNAP_CR_85$date <- as.Date(SNAP_CR_85$date,format="%Y-%m-%d")
SNAP_CR_85

SNAP_YF_45 <- read.csv("BiasCorrection/Geo_BiasCorrection/DataFiles/SNAP_YF45_corrected_monthly.csv")
SNAP_YF_45$date <- as.Date(SNAP_YF_45$date,format="%Y-%m-%d")
SNAP_YF_45
SNAP_YF_85 <- read.csv("BiasCorrection/Geo_BiasCorrection/DataFiles/SNAP_YF85_corrected_monthly.csv")
SNAP_YF_85$date <- as.Date(SNAP_YF_85$date,format="%Y-%m-%d")
SNAP_YF_85
```

# Appending the Temperature Forecasts to the observed data

## Starting with RCP 4.5

### Creating a dataframe with missing values for all other aspects

Determining the columns needed for this

```{r}
CR_ObsData <- read.csv("BiasCorrection/CR_PondAirTemp_mean.csv", header=TRUE)
head(CR_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CR_ObsData <- CR_ObsData %>%
  dplyr::rename(date=Date)
CR_ObsData <- CR_ObsData[,-1]
head(CR_ObsData)

CR_ObsData <- CR_ObsData %>%
  dplyr::select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CR_ObsData

CR_ObsData$date <- as.Date(CR_ObsData$date,format="%Y-%m-%d")
CR_ObsData
```

Columns needed are date, Airport, Air_MonthAvg, Air_MonthMax, Air_MonthMin, Region, TIN, SQR, TIS, CAB, EYS, RHM, BVS, CAN, WDD

#### Copper River Delta

Checking the dimensions of the dataframe

```{r}
SNAP_CR_45
dim(SNAP_CR_45) # 972 rows here, so need to make 972 blank rows
```

Checking the observational data

```{r}
CR_ObsData <- read.csv("BiasCorrection/CR_PondAirTemp_mean.csv", header=TRUE)
head(CR_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CR_ObsData <- CR_ObsData %>%
  dplyr::rename(date=Date)
CR_ObsData <- CR_ObsData[,-1]
head(CR_ObsData)

CR_ObsData <- CR_ObsData %>%
  dplyr::select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CR_ObsData


CR_ObsData$date <- as.Date(CR_ObsData$date,format="%Y-%m-%d")
CR_ObsData
```

Creating an empty dataframe

```{r}
CR_df <- CR_ObsData[0,]
str(CR_df)

attr(CR_df, 'row.names') <- 1:nrow(SNAP_CR_45)

str(CR_df)
```

#### Yakutat Forelands

Checking the dimensions of the dataframe

```{r}
SNAP_YF_45
dim(SNAP_YF_45) # 972 rows here, so need to make 972 blank rows
```

Checking the observational data

```{r}
YF_ObsData <- read.csv("BiasCorrection/YF_PondAirTemp_mean.csv", header=TRUE)
head(YF_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
YF_ObsData <- YF_ObsData %>%
  dplyr::rename(date=Date)
YF_ObsData <- YF_ObsData[,-1]
head(YF_ObsData)

YF_ObsData <- YF_ObsData %>%
  dplyr::select(date, Airport, Air_MonthAvg, std, precision, Region,  MP1, MP3, PL3, PL2, UBP3, MP5, MP8, UBP1, UBP2, PL1, UBP4)
YF_ObsData


YF_ObsData$date <- as.Date(YF_ObsData$date,format="%Y-%m-%d")
YF_ObsData
```

Creating an empty dataframe

```{r}
YF_df <- YF_ObsData[0,]
str(YF_df)

attr(YF_df, 'row.names') <- 1:nrow(SNAP_YF_45)

str(YF_df)
```

### Adding in the data where it is available

#### Copper River Delta

```{r}
CR_df

#values where there is data
CR_df$date <- SNAP_CR_45$date
CR_df$Air_MonthAvg <- SNAP_CR_45$corrected_temp
CR_df$std <- SNAP_CR_45$std_temperature
CR_df$precision <- SNAP_CR_45$precision
CR_df$Region <- rep("CR", nrow(CR_df))

#values where there is not data
CR_df$Airport <- rep(NA, nrow(CR_df))
CR_df$BVS <- rep(NA, nrow(CR_df))
CR_df$CAB <- rep(NA, nrow(CR_df))
CR_df$CAN <- rep(NA, nrow(CR_df))
CR_df$EYS <- rep(NA, nrow(CR_df))
CR_df$RHM <- rep(NA, nrow(CR_df))
CR_df$SQR <- rep(NA, nrow(CR_df))
CR_df$TIN <- rep(NA, nrow(CR_df))
CR_df$TIS <- rep(NA, nrow(CR_df))
CR_df$WDD <- rep(NA, nrow(CR_df))

str(CR_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
library(zoo)  # For as.yearmon function if needed

# Convert CR_df$date to Date format
CR_df$date <- as.Date(as.yearmon(CR_df$date))

# Convert CR_ObsData$date to Date format if not already done
CR_ObsData$date <- as.Date(CR_ObsData$date)

# Combine the dataframes
CR_45Forecast <- rbind(CR_ObsData, CR_df)

# Check the structure of the combined dataframe
str(CR_45Forecast)

#save this as a .csv file
write.csv(CR_45Forecast, file = "CR_45SNAPForecast_MonCorr.csv")

```

#### Yakutat Forelands

Adding in the data where it is available

```{r}
#values where there is data
YF_df$date <- SNAP_YF_45$date
YF_df$Air_MonthAvg <- SNAP_YF_45$corrected_temp
YF_df$std <- SNAP_YF_45$std_temperature
YF_df$precision <- SNAP_YF_45$precision
YF_df$Region <- rep("YF", nrow(YF_df))

#values where there is not data
YF_df$Airport <- rep(NA, nrow(YF_df))
YF_df$MP1<- rep(NA, nrow(YF_df))
YF_df$MP3 <- rep(NA, nrow(YF_df))
YF_df$MP5 <- rep(NA, nrow(YF_df))
YF_df$MP8 <- rep(NA, nrow(YF_df))
YF_df$PL1 <- rep(NA, nrow(YF_df))
YF_df$PL2 <- rep(NA, nrow(YF_df))
YF_df$PL3 <- rep(NA, nrow(YF_df))
YF_df$UBP1 <- rep(NA, nrow(YF_df))
YF_df$UBP2 <- rep(NA, nrow(YF_df))
YF_df$UBP3 <- rep(NA, nrow(YF_df))
YF_df$UBP4 <- rep(NA, nrow(YF_df))

str(YF_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
library(zoo)  # For as.yearmon function if needed

# Convert CR_df$date to Date format
YF_df$date <- as.Date(as.yearmon(YF_df$date))

# Convert CR_ObsData$date to Date format if not already done
YF_ObsData$date <- as.Date(YF_ObsData$date)

# Combine the dataframes
YF_45Forecast <- rbind(YF_ObsData, YF_df)

# Check the structure of the combined dataframe
str(YF_45Forecast)

#save this as a .csv file
write.csv(YF_45Forecast, file = "YF_45SNAPForecast_MonCorr.csv")

```

## Continuing with RCP 8.5

### Creating a dataframe with missing values for all other aspects

Determining the columns needed for this

```{r}
CR_ObsData <- read.csv("BiasCorrection/CR_PondAirTemp_mean.csv", header=TRUE)
head(CR_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CR_ObsData <- CR_ObsData %>%
  dplyr::rename(date=Date)
CR_ObsData <- CR_ObsData[,-1]
head(CR_ObsData)

CR_ObsData <- CR_ObsData %>%
  select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CR_ObsData


CR_ObsData$date <- as.Date(CR_ObsData$date,format="%Y-%m-%d")
CR_ObsData
```

Columns needed are date, Airport, Air_MonthAvg, Air_MonthMax, Air_MonthMin, Region, TIN, SQR, TIS, CAB, EYS, RHM, BVS, CAN, WDD

#### Copper River Delta

Checking the dimensions of the dataframe

```{r}
SNAP_CR_85
dim(SNAP_CR_85) # 972 rows here, so need to make 972 blank rows
```

Checking the observational data

```{r}
CR_ObsData <- read.csv("BiasCorrection/CR_PondAirTemp_mean.csv", header=TRUE)
head(CR_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
CR_ObsData <- CR_ObsData %>%
  dplyr::rename(date=Date)
CR_ObsData <- CR_ObsData[,-1]
head(CR_ObsData)

CR_ObsData <- CR_ObsData %>%
  select(date, Airport, Air_MonthAvg, std, precision, Region, BVS, CAB, CAN, EYS, RHM, SQR, TIN, TIS, WDD)
CR_ObsData

CR_ObsData$date <- as.Date(CR_ObsData$date,format="%Y-%m-%d")
CR_ObsData
```

Creating an empty dataframe

```{r}
CR_df <- CR_ObsData[0,]
str(CR_df)

attr(CR_df, 'row.names') <- 1:nrow(SNAP_CR_85)

str(CR_df)
```

#### Yakutat Forelands

Checking the dimensions of the dataframe

```{r}
SNAP_YF_85
dim(SNAP_YF_85) # 972 rows here, so need to make 972 blank rows
```

Checking the observational data

```{r}
YF_ObsData <- read.csv("BiasCorrection/YF_PondAirTemp_mean.csv", header=TRUE)
head(YF_ObsData)

# remove the first column here and change the column for Date to date
  # Date is a function of some sort so could be the source of errors
YF_ObsData <- YF_ObsData %>%
  dplyr::rename(date=Date)
YF_ObsData <- YF_ObsData[,-1]
head(YF_ObsData)

YF_ObsData <- YF_ObsData %>%
  select(date, Airport, Air_MonthAvg, std, precision, Region,  MP1, MP3, PL3, PL2, UBP3, MP5, MP8, UBP1, UBP2, PL1, UBP4)
YF_ObsData

YF_ObsData$date <- as.Date(YF_ObsData$date,format="%Y-%m-%d")
YF_ObsData
```

Creating an empty dataframe

```{r}
YF_df <- YF_ObsData[0,]
str(YF_df)

attr(YF_df, 'row.names') <- 1:nrow(SNAP_YF_85)

str(YF_df)
```

### Adding in the data where it is available

#### Copper River Delta

```{r}
CR_df

#values where there is data
CR_df$date <- SNAP_CR_85$date
CR_df$Air_MonthAvg <- SNAP_CR_85$corrected_temp
CR_df$std <- SNAP_CR_85$std_temperature
CR_df$precision <- SNAP_CR_85$precision
CR_df$Region <- rep("CR", nrow(CR_df))

#values where there is not data
CR_df$Airport <- rep(NA, nrow(CR_df))
CR_df$BVS <- rep(NA, nrow(CR_df))
CR_df$CAB <- rep(NA, nrow(CR_df))
CR_df$CAN <- rep(NA, nrow(CR_df))
CR_df$EYS <- rep(NA, nrow(CR_df))
CR_df$RHM <- rep(NA, nrow(CR_df))
CR_df$SQR <- rep(NA, nrow(CR_df))
CR_df$TIN <- rep(NA, nrow(CR_df))
CR_df$TIS <- rep(NA, nrow(CR_df))
CR_df$WDD <- rep(NA, nrow(CR_df))

str(CR_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
library(zoo)  # For as.yearmon function if needed

# Convert CR_df$date to Date format
CR_df$date <- as.Date(as.yearmon(CR_df$date))

# Convert CR_ObsData$date to Date format if not already done
CR_ObsData$date <- as.Date(CR_ObsData$date)

# Combine the dataframes
CR_85Forecast <- rbind(CR_ObsData, CR_df)

# Check the structure of the combined dataframe
str(CR_85Forecast)

#save this as a .csv file
write.csv(CR_85Forecast, file = "CR_85SNAPForecast_MonCorr.csv")

```

#### Yakutat Forelands

Adding in the data where it is available

```{r}
YF_df

#values where there is data
YF_df$date <- SNAP_YF_85$date
YF_df$Air_MonthAvg <- SNAP_YF_85$corrected_temp
YF_df$std <- SNAP_YF_85$std_temperature
YF_df$precision <- SNAP_YF_85$precision
YF_df$Region <- rep("YF", nrow(YF_df))

#values where there is not data
YF_df$Airport <- rep(NA, nrow(YF_df))
YF_df$MP1<- rep(NA, nrow(YF_df))
YF_df$MP3 <- rep(NA, nrow(YF_df))
YF_df$MP5 <- rep(NA, nrow(YF_df))
YF_df$MP8 <- rep(NA, nrow(YF_df))
YF_df$PL1 <- rep(NA, nrow(YF_df))
YF_df$PL2 <- rep(NA, nrow(YF_df))
YF_df$PL3 <- rep(NA, nrow(YF_df))
YF_df$UBP1 <- rep(NA, nrow(YF_df))
YF_df$UBP2 <- rep(NA, nrow(YF_df))
YF_df$UBP3 <- rep(NA, nrow(YF_df))
YF_df$UBP4 <- rep(NA, nrow(YF_df))

str(YF_df)
```

Appending this new dataframe "df" to the CR_ObsData dataframe

```{r}
library(zoo)  # For as.yearmon function if needed

# Convert CR_df$date to Date format
YF_df$date <- as.Date(as.yearmon(YF_df$date))

# Convert CR_ObsData$date to Date format if not already done
YF_ObsData$date <- as.Date(YF_ObsData$date)

# Combine the dataframes
YF_85Forecast <- rbind(YF_ObsData, YF_df)

# Check the structure of the combined dataframe
str(YF_85Forecast)

#save this as a .csv file
write.csv(YF_85Forecast, file = "YF_85SNAPForecast_MonCorr.csv")

```

# Attemping a quick plot of this data for both CR and YF

## RCP 4.5

### Copper River Delta

```{r}
CR_45Forecast 

CR_45Temps <- ggplot(data = CR_45Forecast, aes(x = date)) +
  geom_line(aes(y = Air_MonthAvg)) +  # Line plot for average temperature
  geom_ribbon(aes(ymin = Air_MonthAvg - std, ymax = Air_MonthAvg + std),  # Ribbon for uncertainty
              alpha = 0.5, fill = "blue") +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Copper River Delta Temperature - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
CR_45Temps

ggsave("CR_45Temps.jpeg", plot = CR_45Temps, width = 8, height = 6)
```

### Yakutat Forelands

```{r}
YF_45Forecast 

YF_45Temps<- ggplot(data = YF_45Forecast, mapping = aes(x = date)) +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
   geom_ribbon(aes(ymin = Air_MonthAvg - std, ymax = Air_MonthAvg + std),  # Ribbon for uncertainty
              alpha = 0.5, fill = "blue") +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Yakutat Forelands Temperature - RCP 4.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
YF_45Temps

ggsave("YF_45Temps.jpeg", plot = YF_45Temps, width = 8, height = 6)
```

## RCP 8.5

### Copper River Delta

```{r}
CR_85Forecast 

CR_85Temps<- ggplot(data = CR_85Forecast, mapping = aes(x = date)) +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
   geom_ribbon(aes(ymin = Air_MonthAvg - std, ymax = Air_MonthAvg + std),  # Ribbon for uncertainty
              alpha = 0.5, fill = "blue") +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Copper River Delta Temperature - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
CR_85Temps

ggsave("CR_85Temps.jpeg", plot = CR_85Temps, width = 8, height = 6)
```

### Yakutat Forelands

```{r}
YF_85Forecast 

YF_85Temps<- ggplot(data = YF_85Forecast, mapping = aes(x = date)) +
  geom_line(mapping = aes(y = Air_MonthAvg)) +
   geom_ribbon(aes(ymin = Air_MonthAvg - std, ymax = Air_MonthAvg + std),  # Ribbon for uncertainty
              alpha = 0.5, fill = "blue") +
  xlab("Year") +
  ylab("Mean Temperature (C)") +
  ggtitle("Yakutat Forelands Temperature - RCP 8.5") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
YF_85Temps

ggsave("YF_85Temps.jpeg", plot = YF_85Temps, width = 8, height = 6)
```
